### 2023-11-16 Paper_code-Database_CT.Rmd
worked on coactivation in tumors

---
title: "Code used to generate paper results : Database comparison and CT characteristics"
author: "Julie Devis"
date: "2023-08-28"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

# Loading data and packages

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(Vennerable)
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(UpSetR)
library(ComplexHeatmap)
library(circlize)
library(SingleCellExperiment)
library(org.Hs.eg.db)
library(clusterProfiler)
library(msigdbr)
library(DOSE)
library(BiocParallel)
library(patchwork)
library(Biostrings)
```

Gene names/synonyms required for databases cleaning

```{r ensembl_synonym, warning=FALSE, message=FALSE}
ensembl <- biomaRt::useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
attributes_vector <- c("ensembl_gene_id", "external_gene_name",
                       "external_synonym", "gene_biotype",
                       "chromosome_name", "band", "start_position", "end_position",
                       "strand")
ensembl_gene_synonym <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))

ensembl_gene_synonym <- ensembl_gene_synonym %>%
  mutate(external_synonym = sub(pattern = "ORF", external_synonym, 
                                replacement = "orf"))

attributes_vector <- c("ensembl_gene_id", "external_gene_name")
ensembl_gene_names <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))
```

```{r load_data, warning=FALSE, message=FALSE}
GTEX_data <- CTdata::GTEX_data()
normal_tissues_multimapping_data <- CTdata::normal_tissues_multimapping_data()
CCLE_data <- CTdata::CCLE_data()
TCGA_TPM <- CTdata::TCGA_TPM()
DAC_treated_cells_multimapping <- CTdata::DAC_treated_cells_multimapping()
testis_sce <- CTdata::testis_sce()
scRNAseq_HPA <- CTdata::scRNAseq_HPA()
CT_mean_methylation_in_tissues <- CTdata::CT_mean_methylation_in_tissues()
```

Common figures parameters

```{r fig_param}
legends_param <- list(
  labels_gp = gpar(col = "black", fontsize = 6),
  title_gp = gpar(col = "black", fontsize = 6),
  row_names_gp = gpar(fontsize = 4),
  annotation_name_side = "left")

legend_colors <- c("#5E4FA2", "#3288BD", "#66C2A5", "#ABDDA4", "#E6F598",
                   "#FFFFBF", "#FEE08B", "#FDAE61", "#F46D43", "#D53E4F",
                   "#9E0142")
chr_colors <- c("X-linked" = "deeppink", "Not X" = "royalblue1")

meth_colors <- c("Methylation" = "lightgreen", "Not methylation" = "gray")
```

# Database comparison

## Lists cleaning

CT lists from other databases have been checked (using GTEx and our 
`GTEx_expression()` funtion and GeneCards) in order to remove duplicated
gene names or deprecated ones and allow comparison between databases.

### CTdatabase

Online list copied in a csv file, several lists exist so we combined them.

We checked gene names that were a concetanation of two genes (choice using
biomaRt synonyms to get the official one), checked which ones had the right names,
removed duplicated genes, verified lost genes and added back those that should 
be there.

```{r CTdatabase_cleaning, message = FALSE}
CTdatabase <- read_delim("data/CTdatabase1.csv", delim = ";", 
                         escape_double = FALSE, trim_ws = TRUE)
colnames(CTdatabase) <- c("Family", "Gene_Name", "Chromosomal_localization",
                          "CT_identifier")
CTdatabase_bis <- read_csv2("data/CTdatabase2.csv")
CTdatabase <- left_join(CTdatabase, CTdatabase_bis, 
                        by = c("Gene_Name" = "Gene_Symbol"))

CTdatabase_single <- CTdatabase %>%
  mutate(Gene_Name = sub(pattern = "/.*$", Gene_Name, replacement = ""))
CTdatabase_single <- CTdatabase_single %>%
  mutate(Gene_Name = sub(pattern = ",.*$", Gene_Name, replacement = ""))

CTdatabase_official_names <- 
  unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, 
                       external_gene_name)) %>%
  filter(external_gene_name %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA)
CTdatabase_synonym <- 
  ensembl_gene_synonym %>%
  filter(external_synonym %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_synonym) %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym)
CTdatabase_cleaned <- 
  rbind(CTdatabase_official_names, CTdatabase_synonym) %>% 
  left_join(CTdatabase_single)

duplicated_genes <- CTdatabase_cleaned$Gene_Name[duplicated(CTdatabase_cleaned$Gene_Name)]
bad_ids <- ensembl_gene_synonym %>%
  filter(external_gene_name %in% duplicated_genes | external_synonym %in% duplicated_genes) %>%
  filter(chromosome_name %in% grep(pattern = "H", x = chromosome_name, value = TRUE)) %>%
  pull(ensembl_gene_id)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  dplyr::filter(!ensembl_gene_id %in% bad_ids)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  filter(!ensembl_gene_id == "ENSG00000052126")
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!(ensembl_gene_id == "ENSG00000183305" & Gene_Name == "MAGEA2"))
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!ensembl_gene_id == "ENSG00000204648")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CSAG3B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CSAG2", "external_synonym"] <- "CSAG3B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT45A4")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CT45A3", "external_synonym"] <- "CT45A4"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "LAGE-1b")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAG2", "external_synonym"] <- "LAGE-1b"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT16.2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "PAGE5", "external_synonym"] <- "CT16.2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXB2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXB1", "external_synonym"] <- "SPANXB2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXE")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXD", "external_synonym"] <- "SPANXE"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1C")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1D")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1E")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE2B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "XAGE2", "external_synonym"] <- "XAGE2B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CTAGE-2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAGE1", "external_synonym"] <- "CTAGE-2"

CTdatabase_cleaned <- ensembl_gene_synonym %>%
  mutate(Gene_Name = external_synonym) %>%
  filter(external_synonym == "CXorf61") %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym) %>%
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "Cxorf61", 
                    c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name)) %>%
  filter(external_gene_name == "CCNA1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "cyclin A1", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "GOLGA6L2") %>%
  filter(ensembl_gene_id == "ENSG00000174450") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "GOLGAGL2 FA", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LYPD6B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC130576", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT62") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC196993", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT75") %>%
  filter(ensembl_gene_id == "ENSG00000291155") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC440934", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LINC01192") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC647107", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "TSPY1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC728137", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "SSX2B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "SSX2b", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
```

### Jamin's list

Excel file coming from supplemental data.

```{r Jamin_clean, message = FALSE}
Jamin_core_CT <- read_excel("data/Jamin_core_CT.xlsx")
Jamin_core_CT[Jamin_core_CT$Gene == "KIAA1211", "Gene"] <- "CRACD"
Jamin_core_CT[Jamin_core_CT$Gene == "CXorf67", "Gene"] <- "EZHIP"
```

### Wang's CTatlas

Excel file coming from supplemental data.

```{r Wang_cleaning, message = FALSE}
Wang_CT <- read_excel("data/Wang_Suppl_Data_3.xlsx", 
    sheet = "Supplementary Data 3B", skip = 1)
colnames(Wang_CT)[1] <- "ensembl_gene_id"

Wang_CT <- ensembl_gene_names %>% 
  filter(ensembl_gene_id %in% Wang_CT$ensembl_gene_id) %>%
  right_join(Wang_CT)

Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000181013", "external_gene_name"] <- "C17orf47"
Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000204293", "external_gene_name"] <- "OR8B2"
Wang_CT[Wang_CT$external_gene_name == "", "external_gene_name"] <- "RNASE11"
Wang_CT[Wang_CT$external_gene_name == "CHCT1", "external_gene_name"] <- "C17orf64"
Wang_CT[Wang_CT$external_gene_name == "PRSS40A", "external_gene_name"] <- "TISP43"
Wang_CT[Wang_CT$external_gene_name == "TEX56P", "external_gene_name"] <- "C6orf201"
Wang_CT[Wang_CT$external_gene_name == "SLC25A51P4", "external_gene_name"] <- "RP11-113D6.10"
Wang_CT[Wang_CT$external_gene_name == "TCP10L3", "external_gene_name"] <- "TCP10"
Wang_CT[Wang_CT$external_gene_name == "SCAND3", "external_gene_name"] <- "ZBED9"
```

### Carter's list

```{r Carter_clean, message = FALSE}
Carter_CT_list <- read_table("data/Carter_CT_list.txt", skip = 1)
Carter_CT <- Carter_CT_list[Carter_CT_list$CT_Expression,]

Carter_CT[Carter_CT$Gene == "ENSG00000261649", "Gene_Name"] <- "GOLGA6L7"
Carter_CT[Carter_CT$Gene == "ENSG00000239620", "Gene_Name"] <- "PRR20G"
Carter_CT[Carter_CT$Gene == "ENSG00000168148", "Gene_Name"] <- "H3-4"
Carter_CT[Carter_CT$Gene == "ENSG00000204296", "Gene_Name"] <- "TSBP1"
Carter_CT[Carter_CT$Gene == "ENSG00000180219", "Gene_Name"] <- "GARIN6"
Carter_CT[Carter_CT$Gene == "ENSG00000172717", "Gene_Name"] <- "GARIN2"
Carter_CT[Carter_CT$Gene == "ENSG00000174015", "Gene_Name"] <- "CBY2"
Carter_CT[Carter_CT$Gene == "ENSG00000224960", "Gene_Name"] <- "PPP4R3C"
```

## CTexploreR data for selection pipeline

To characterise the differences between our database and other, we need the 
category we created in CTexploreR.

Hereunder is what we used for our selection pipeline (coming from 
`make_CT_list.R` in `CTdata`), mainly how we created the data.

```{r make_CT, message = FALSE, warning = FALSE}
# GTEX data with the tissue specificity category determined
all_genes <- as_tibble(rowData(GTEX_data), rownames = "ensembl_gene_id")
all_genes$TPM_testis <- assay(GTEX_data)[, "Testis"]

# Add multimapping_analysis assessing testis-specificity of genes "lowly_expressed"

all_genes <- all_genes %>%
  left_join(as_tibble(rowData(normal_tissues_multimapping_data), 
                      rownames = "ensembl_gene_id"))

# Summarise both specificity

all_genes <- all_genes %>%
  mutate(testis_specificity = case_when(
    GTEX_category == "testis_specific" |
      multimapping_analysis == "testis_specific" ~ "testis_specific",
    GTEX_category == "testis_preferential" ~ "testis_preferential"))

# Add testis scRNA seq highest expressed cell type

all_genes <- all_genes %>%
  left_join(as_tibble(rowData(testis_sce)) %>%
              dplyr::select(external_gene_name, testis_cell_type))

# Add higher in somatic scRNAseq data of normal tissues from the Human Protein Atlas

all_genes <- all_genes %>%
  left_join(as_tibble(rowData(scRNAseq_HPA), rownames = "ensembl_gene_id") %>%
              dplyr::select(ensembl_gene_id, external_gene_name,
                            Higher_in_somatic_cell_type))

# CCLE database analysis added

all_genes <- all_genes %>%
  left_join(as_tibble(rowData(CCLE_data), rownames = "ensembl_gene_id"))
all_genes[is.na(all_genes$CCLE_category), "CCLE_category"] <- "not_in_CCLE"

# TCGA database analysis added

all_genes <- all_genes %>%
  left_join(as_tibble(rowData(TCGA_TPM), rownames = "ensembl_gene_id") %>%
              dplyr::select(ensembl_gene_id, external_gene_name, percent_pos_tum,
                            percent_neg_tum, max_TPM_in_TCGA, TCGA_category))

all_genes[all_genes$lowly_expressed_in_GTEX == TRUE &
            all_genes$multimapping_analysis == "testis_specific",
          "TCGA_category"] <- "multimapping_issue"

# Add DAC analysis
induced <- as_tibble(rowData(DAC_treated_cells_multimapping), 
                     rownames = "ensembl_gene_id") %>%
  filter(induced == TRUE) %>%
  pull(external_gene_name)

all_genes <- all_genes %>%
  mutate(DAC_induced = case_when(external_gene_name %in% induced ~ TRUE,
                                 !external_gene_name %in% induced ~ FALSE))

# Add the most biologically relevant transcript (canonical and coordinates)

ensembl <- biomaRt::useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
attributes_vector <- c("ensembl_gene_id",
                       "external_gene_name",
                       "ensembl_transcript_id",
                       "external_transcript_name",
                       "chromosome_name",
                       "strand",
                       "transcript_start",
                       "transcript_end",
                       "transcription_start_site",
                       "transcript_length",
                       "transcript_biotype",
                       "transcript_is_canonical")
transcripts_infos <- as_tibble(biomaRt::getBM(attributes = attributes_vector,
                                              mart = ensembl))
canonical_transcripts <- transcripts_infos %>%
  filter(transcript_is_canonical == 1) %>%
  filter(chromosome_name %in% c(1:22, "X", "Y", "MT")) %>%
  filter(transcript_biotype == "protein_coding" | transcript_biotype == "lncRNA")
all_genes <- all_genes %>%
  left_join(canonical_transcripts %>%
              dplyr::select(ensembl_gene_id,
                            external_transcript_name, ensembl_transcript_id,
                            chromosome_name, strand, transcript_start,
                            transcript_end, transcription_start_site,
                            transcript_length, transcript_biotype))

```

From there, we filtered based on the testis_specificity ("testis_specific" or 
"testis_preferential"), testis_cell_type (not "Macrophage", "Endothelial",
"Myoid", "Sertoli", "Leydig"), Higher_in_somatic_cell_type (TRUE), 
CCLE_category ("activated") and TCGA_category ("activated" or 
"multimapping_issue") to have our CT genes. Then the most relevant 
transcript was validated manually (and check that there is a transcript 
activated in testis and the same is activated in tumor).

## CTexploreR VS CTdatabase

```{r CTdatabase_ours}
CTdatabase_ours <- Venn(list(CTdatabase = CTdatabase_cleaned$external_gene_name,
                             CTexploreR = CT_genes$external_gene_name))
gp <- VennThemes(compute.Venn(CTdatabase_ours))
gp[["Face"]][["11"]]$fill <-  "mistyrose"
gp[["Face"]][["01"]]$fill <-  "darkseagreen1"
gp[["Face"]][["10"]]$fill <-  "lightsteelblue1"
gp[["Set"]][["Set1"]]$col <-  "paleturquoise4"
gp[["Set"]][["Set2"]]$col <-  "darkseagreen4"
gp[["SetText"]][["Set1"]]$col <-  "paleturquoise4"
gp[["SetText"]][["Set2"]]$col <-  "darkseagreen4"
plot(CTdatabase_ours, gp = gp)

```

Lost genes analysis

```{r CTdatabase_lost}
CTdatabase_lost <- all_genes %>%
  filter(external_gene_name %in% CTdatabase_ours@IntersectionSets[["10"]])

# 10 Genes are lost because not in any  database

CTdatabase_lost[is.na(CTdatabase_lost$testis_specificity), ]$testis_specificity <- "not_specific"
table(CTdatabase_lost$testis_specificity)

table(CTdatabase_lost$testis_cell_type)

table(CTdatabase_lost$Higher_in_somatic_cell_type)

table(CTdatabase_lost$TCGA_category)

table(CTdatabase_lost$CCLE_category)
```

## CTexploreR VS omics databases
```{r upset_plot}
core_ours <- Venn(list(Jamin = Jamin_core_CT$Gene, 
                       CTexploreR = CT_genes$external_gene_name))

Wang_ours <- Venn(list(Wang = Wang_CT$external_gene_name, 
                       CTexploreR = CT_genes$external_gene_name))

Carter_ours <- Venn(list(Carter_CT = Carter_CT$Gene_Name, 
                         CTexploreR = CT_genes$external_gene_name))
gene_list <- list(Jamin = Jamin_core_CT$Gene, 
                  CTdatabase = CTdatabase_cleaned$external_gene_name, 
                  CTatlas = Wang_CT$external_gene_name, 
                  CTexploreR = CT_genes$external_gene_name, 
                  Carter = Carter_CT$Gene_Name)

upset_omics <- fromList(gene_list[-2])
upset(upset_omics)
```

Lost genes analysis

```{r Jamin_lost}
Jamin_lost <- all_genes %>%
  filter(external_gene_name %in% core_ours@IntersectionSets[["10"]])

Jamin_lost[is.na(Jamin_lost$testis_specificity), ]$testis_specificity <- "not_specific"
table(Jamin_lost$testis_specificity)

table(Jamin_lost$testis_cell_type)

table(Jamin_lost$Higher_in_somatic_cell_type)

table(Jamin_lost$TCGA_category)

table(Jamin_lost$CCLE_category)
```

```{r Wang_lost}
Wang_lost <- all_genes %>%
  filter(external_gene_name %in% Wang_ours@IntersectionSets[["10"]])

Wang_lost[is.na(Wang_lost$testis_specificity), ]$testis_specificity <- "not_specific"
table(Wang_lost$testis_specificity)

table(Wang_lost$testis_cell_type)

table(Wang_lost$Higher_in_somatic_cell_type)

table(Wang_lost$TCGA_category)

table(Wang_lost$CCLE_category)

```

```{r Carter_lost}
Carter_lost <- all_genes %>%
  filter(external_gene_name %in% Carter_ours@IntersectionSets[["10"]])

Carter_lost[is.na(Carter_lost$testis_specificity), ]$testis_specificity <- "not_specific"
table(Carter_lost$testis_specificity)

table(Carter_lost$testis_cell_type)

table(Carter_lost$Higher_in_somatic_cell_type)

table(Carter_lost$TCGA_category)

table(Carter_lost$CCLE_category)
```

## Characterisation of differences with all databases

```{r all_databases}
common <- unique(c(core_ours@IntersectionSets[["11"]], 
                   CTdatabase_ours@IntersectionSets[["11"]], 
                   Wang_ours@IntersectionSets[["11"]], 
                   Carter_ours@IntersectionSets[["11"]]))

length(common)

length(common)/dim(CT_genes)[1] * 100

lost_list <- unique(c(core_ours@IntersectionSets[["10"]], 
                 CTdatabase_ours@IntersectionSets[["10"]],
                 Wang_ours@IntersectionSets[["10"]],
                 Carter_ours@IntersectionSets[["10"]]))

lost <- all_genes %>%
  filter(external_gene_name %in% lost_list)

not_specific <- filter(lost, is.na(testis_specificity))

GTEX_expression(not_specific$external_gene_name, units = "log_TPM")

somatic_testis <- filter(lost, testis_cell_type %in% c( "Macrophage", 
                                                        "Endothelial", "Myoid",
                                                        "Sertoli", "Leydig"))

testis_expression(somatic_testis$external_gene_name, cells = "all")

somatic_tissue <- filter(lost, Higher_in_somatic_cell_type == TRUE)

HPA_cell_type_expression(somatic_tissue$external_gene_name)

not_TCGA_activated <- filter(lost, TCGA_category != "activated" & 
                               TCGA_category != "multimapping_issue")

TCGA_expression(not_TCGA_activated$external_gene_name,
                tumor = "all",
                units = "log_TPM")

not_CCLE_activated <- filter(lost, CCLE_category  != "activated")

CCLE_expression(not_CCLE_activated$external_gene_name,
                  type = c("lung", "skin", "colorectal",
                           "gastric", "breast", "head_and_neck"),
                units = "log_TPM")

transcript_prob <- lost %>% 
  filter(testis_specificity == "testis_specific" |
           testis_specificity == "testis_preferential") %>%
  filter(TCGA_category == "activated" | TCGA_category == "multimapping_issue") %>% 
  filter(CCLE_category == "activated") %>% 
  dim()
```

We have lost `r length(lost_list)` genes in total. Among them, 
`r dim(not_specific)[1]/length(lost_list)*100`% are not considered testis 
specific, `r dim(somatic_testis)[1]/length(lost_list)*100`% are expressed in 
testis somatic cells, `r dim(somatic_tissue)[1]/length(lost_list)*100`% are 
expressed in somatic cells, `r dim(not_TCGA_activated)[1]/length(lost_list)*100`% 
are not activated in TCGA samples, 
`r dim(not_CCLE_activated)[1]/length(lost_list)*100`% are not activated in CCLE 
cell lines and `r transcript_prob[1]/length(lost_list)*100`% is lost due to 
transcripts problems.

What about new genes in CTexploreR

```{r, warning = FALSE}
new <- CT_genes %>% 
  filter(!external_gene_name%in%common)

new

table(new$testis_specificity)
table(new$X_linked, new$regulated_by_methylation)
TCGA_expression(tumor = "all", genes = new$external_gene_name, 
                units = "log_TPM")
TCGA_expression(tumor = "all", 
                genes = filter(new, X_linked & regulated_by_methylation)$external_gene_name, 
                units = "log_TPM")
```
There are `r dim(new)[1]` new CT genes in CTexploreR. These are mainly testis
specific and on autosomes. Regulation by methylation is not the majority of 
them. This makes sens as other databases have been focused on regulation by 
methylation, so as here we don't segregate with that we can find these. There 
is only 11 new "major" CT that are on the X chromosome and regulated by 
methylation. CT45 are not that new.

Expression in tumours doesn't strike that much.

# CT genes characteristics

```{r basic}
table(CT_genes$testis_specificity)
table(CT_genes$transcript_biotype)
```

Most genes are testis specific 
(`r table(CT_genes$testis_specificity)[[2]]/dim(CT_genes)[1]*100`%).
Most genes are mainly protein coding genes 
(`r table(CT_genes$transcript_biotype)[[3]]/dim(CT_genes)[1]*100`%).

## X chromosome and regulated by methylation

In CTexploreR, genes have been characterised as regulated by methylation or not.

```{r X_meth}
table(CT_genes$X_linked)
table(CT_genes$regulated_by_methylation)
table(CT_genes$X_linked, CT_genes$testis_specificity)
table(CT_genes$X_linked, CT_genes$regulated_by_methylation)
```

Genes are enriched on the X chromosome 
(`r table(CT_genes$X_linked)[[2]]/dim(CT_genes)[1]*100`%). Also, 
`r table(CT_genes$regulated_by_methylation)[[2]]` genes have been flagged as 
regulated by methylation 
(`r table(CT_genes$regulated_by_methylation)[[2]]/dim(CT_genes)[1]*100`%). 
It is interesting to study the link between these two characteristics.

On the chromosome X, there is a clear enrichment of CT genes regulated by 
methylation (83/93 chrX genes or 83/160 genes regulated by methylation).

Let's check that with a statistical test, I here want to see if, on there is an 
enrichment of genes regulated by methylation on the X chromosome. I need to do a
Pearson Chi square test (to know if observed proportion differ from expected 
proportion). It is a statistical method to determine if two categorical variables
have a significant correlation between them. I can directly put a matrix 
(my table) in the function

```{r chisq}
chisq.test(table(CT_genes$X_linked, CT_genes$regulated_by_methylation))
```

```{r X_meth_fig}
CT_genes$chr_factor <- factor(CT_genes$chr,
                       levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9",
                                  "10", "11", "12", "13", "14", "15", "16", "17", 
                                  "18", "19", "20", "21", "22", "X", "Y"))

CT_genes %>% 
  mutate(regulated_by_methylation = ifelse(regulated_by_methylation,
                                           "Regulated by methylation",
                                           "Not regulated by methylation")) %>% 
  ggplot(aes(x = chr_factor, fill = testis_specificity, color = X_linked)) +
  geom_bar(stat = 'count') +
  scale_fill_manual(values = c("goldenrod1", "mediumpurple1")) +
  scale_color_manual(values = c("royalblue1", "deeppink"), guide = "none") +
  facet_wrap(~ regulated_by_methylation, ncol = 2) +
  theme_bw() +
  xlab("Chromosome") +
  ylab("Number of genes") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top",
        axis.title = element_text(size = 10, color = "gray10"))
```

There is indeed a significative link between regulation by methylation and being
on the X chromosome. There is thus an enrichment of CT genes regulated by 
methylation on the X chromosome (and inversely).

## Tumour and methylation

Using CTexploreR functions, we can explore all CT genes or focus on some 
potential targets.

### All CT

For these heatmaps, the code comes from the function but has been copies to add 
some annotations.

```{r common_annot}
chr_mat <- as.matrix(CT_genes$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes$external_gene_name
row_ha_chr <- rowAnnotation(chr_factor = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr_factor = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes$regulated_by_methylation, CT_genes$X_linked)
```

```{r TCGA_heatmap, message = FALSE}
database <- TCGA_TPM
database$tumor <- sub(pattern = 'TCGA-', x = database$project_id, '')
database$type <- "Tumor"
database$type[database$shortLetterCode == "NT"] <- "Peritumoral"
database <- database[, order(database$tumor, database$type)]

genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]
database <- database[, database$type == "Tumor"]

column_ha_tumor <- HeatmapAnnotation(
  Tumor = database$tumor,
  border = TRUE,
  col = list(Tumor = c("BRCA" = "midnightblue", "COAD" = "darkorchid2",
                 "ESCA" = "gold", "HNSC" = "deeppink2",
                 "LUAD" = "seagreen", "LUSC" = "seagreen3",
                 "SKCM" = "red3")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

split_by <- factor(database$tumor)
col_annot <- column_ha_tumor

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

Heatmap(mat[, , drop = FALSE],
        name = "logTPM",
        column_title = paste0("Expression in TCGA samples (all)"),
        column_split = split_by,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                         legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = col_annot,
        left_annotation = left_annot)

```

```{r CCLE_heatmap}
database <- CCLE_data
database$type <- tolower(database$type)
genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

df_col <- data.frame("cell_line" = colData(database)$cell_line_name,
                     "type" = as.factor(colData(database)$type))
rownames(df_col) <- rownames(colData(database))
df_col <- df_col[order(df_col$type), ]

column_ha_type <- HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param,
  col = list(type = c("lung" = "seagreen3", "skin" = "red3",
                 "bile_duct" = "mediumpurple1", "bladder" = "mistyrose2",
                 "colorectal" = "plum", "lymphoma" = "steelblue1",
                 "uterine" = "darkorange4", "myeloma" = "turquoise3", 
                 "kidney" = "thistle4",
                 "pancreatic" = "darkmagenta", "brain" = "palegreen2",
                 "gastric" = "wheat3", "breast" = "midnightblue",
                 "bone" = "sienna1", "head_and_neck" = "deeppink2",
                 "ovarian" = "tan3", "sarcoma" = "lightcoral",
                 "leukemia" = "steelblue4", "esophageal"= "khaki",
                 "neuroblastoma" = "olivedrab1")))

Heatmap(mat[, rownames(df_col), drop = FALSE],
        name = "logTPM",
        column_title = "Gene Expression in tumor cell lines (CCLE)",
        column_split = factor(df_col$type),
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                          legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = c(column_ha_type),
        left_annotation = left_annot)
```

Next graph was removed from paper

```{r all_meth}
genes <- CT_genes$external_gene_name
database <- CT_mean_methylation_in_tissues[rownames(CT_mean_methylation_in_tissues) %in% genes]

mat <- na.omit(assay(database))
clustering_option <- TRUE

row_ha_chr_meth <- rowAnnotation(chr = chr_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

row_ha_reg_meth <- rowAnnotation(regulation = regulation_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot_meth <- c(row_ha_chr_meth, row_ha_reg_meth, gap = unit(1, "mm"))

split_meth <- data.frame(filter(CT_genes, external_gene_name %in% rownames(mat))$regulated_by_methylation, 
                    filter(CT_genes, external_gene_name %in% rownames(mat))$X_linked)

Heatmap(mat,
        column_title = 'Promoter mean methylation level by tissue',
        name = 'Meth',
        col = colorRamp2(c(1:100),
                         colorRampPalette(c("moccasin","dodgerblue4"))(100)),
        na_col = "gray80",
        cluster_rows = clustering_option,
        cluster_columns = FALSE,
        row_split = split_meth,
        row_title_gp = gpar(fontsize = 0),
        show_row_names = TRUE,
        show_heatmap_legend = TRUE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 3),
        column_names_gp = gpar(fontsize = 8),
        column_names_side = "bottom",
        row_names_side = "right",
        left_annotation = left_annot_meth)

```

### Coactivation in tumours

We know that CT genes that are regulated by methylation are activated in 
tumours due to methylation. However, those that are not it is more blurry. It 
is interesting to look at their expression to see if there are groups of genes 
that are activated together in tumours which would suggest another activation :
a transcription factor ? a gene expresion program from meiosis ? timing of 
expression ? function ?

```{r not_reg}
not_regulated <- filter(CT_genes, !regulated_by_methylation)$external_gene_name
```

As they can blur the clustering, I sometimes remove genes that are in the 
background and  those that are expressed a lot.

- BRCA :

```{r BRCA}
TCGA_expression(filter(CT_genes, !regulated_by_methylation)$external_gene_name,
                tumor = "BRCA", units = "log_TPM")
# Removing genes to see clearer

BRCA_genes <- rowData(TCGA_TPM) %>% 
  as_tibble() %>% 
  filter(external_gene_name %in% not_regulated) %>% 
  filter(percent_pos_BRCA > 0.01 & percent_pos_BRCA < 30) %>% 
  pull(external_gene_name)

BRCA_genes_ordered <- 
  BRCA_genes[row_order(draw(TCGA_expression(BRCA_genes, tumor = "BRCA", 
                                            units = "log_TPM")))]

BRCA_cluster1 <- BRCA_genes_ordered[c(which(BRCA_genes_ordered == "AKAP14"):
                       which(BRCA_genes_ordered == "C11orf97"))]
BRCA_cluster2 <- BRCA_genes_ordered[c(which(BRCA_genes_ordered == "GTF2I-AS1"):
                       which(BRCA_genes_ordered == "SLIT3-AS1"))]

```

We can see two small clusters.

- LUAD :

```{r LUAD}
TCGA_expression(filter(CT_genes, !regulated_by_methylation)$external_gene_name,
                tumor = "LUAD", units = "log_TPM")
# Removing genes to see clearer
LUAD_genes <- rowData(TCGA_TPM) %>% 
  as_tibble() %>% 
  filter(external_gene_name %in% not_regulated) %>% 
  filter(percent_pos_LUAD > 0.01 & percent_pos_LUAD < 30) %>% 
  pull(external_gene_name)

LUAD_genes_ordered <- 
  LUAD_genes[row_order(draw(TCGA_expression(LUAD_genes, tumor = "LUAD", 
                                            units = "log_TPM")))]

LUAD_cluster1 <- c("TSPY1", "TSPY2")
LUAD_cluster2 <- c("SPANXC", "SPANXB1",  "SPANXD")
LUAD_cluster3 <- LUAD_genes_ordered[c(1:22)]
LUAD_cluster4 <- LUAD_genes_ordered[c(which(LUAD_genes_ordered == "GTF2I-AS1"):
                       which(LUAD_genes_ordered == "SLIT3-AS1"))]
```

22 first genes are clearly quite clustering, there is also a small SPANX cluster,
a small TSPY cluster ans a bigger one around the middle of the graph.

- LUSC :

```{r LUSC}
TCGA_expression(filter(CT_genes, !regulated_by_methylation)$external_gene_name,
                tumor = "LUSC", units = "log_TPM")
# Removing genes to see clearer

LUSC_genes <- rowData(TCGA_TPM) %>% 
  as_tibble() %>% 
  filter(external_gene_name %in% not_regulated) %>% 
  filter(percent_pos_LUSC > 0.01 & percent_pos_LUSC < 30) %>% 
  pull(external_gene_name)

LUSC_genes_ordered <- 
  LUSC_genes[row_order(draw(TCGA_expression(LUSC_genes, tumor = "LUSC", 
                                            units = "log_TPM")))]

LUSC_cluster1 <- c("TSPY1", "TSPY2",  "TSPY9", "TSPY10")
LUSC_cluster2 <- c("SPANXC", "SPANXB1",  "SPANXD")
LUSC_cluster3 <- LUSC_genes_ordered[c(1:15)]

```

We can see a cluster at the top and two smaller ones corresponding to SPANX and
TSPY genes. As well as quite a big cluster with the 15 first lines

- SKCM :

```{r SKCM}
SKCM_genes <- rowData(TCGA_TPM) %>% 
  as_tibble() %>% 
  filter(external_gene_name %in% not_regulated) %>% 
  filter(percent_pos_SKCM > 0.01 & percent_pos_SKCM < 30) %>% 
  pull(external_gene_name)

TCGA_expression(filter(CT_genes, !regulated_by_methylation)$external_gene_name,
                tumor = "SKCM", units = "log_TPM")
# Lots of genes activated almost everywhere or nowhere that can mask clusters
TCGA_expression(SKCM_genes,
                tumor = "SKCM", units = "log_TPM")

SKCM_cluster <- c("TSPY1", "TSPY2", "TSPY3", "TSPY9", "TSPY10")
```

It is clear that there is a TSPY cluster again.

- Comparison

TSPY and SPANX genes are found reactivated/coactivated in different types of 
tumours. Let's compare the bigger groups found in LUAD, LUSC and BRCA

```{r upste_clusters}
upset(fromList(list(LUSC = LUSC_cluster3,
               LUAD1 = LUAD_cluster3,
               LUAD2 = LUAD_cluster4,
               BRCA1 = BRCA_cluster1,
               BRCA2 = BRCA_cluster2)))
```

There is quite a lot of similarities here between the clusters.

Let's have a look at the timings of expression.

```{r timing_cluster, message = FALSE}
testis_expression(c("TSPY1", "TSPY2", "TSPY3", "TSPY9", "TSPY10"), 
                  cells = "all")
testis_expression(c("SPANXC", "SPANXB1",  "SPANXD"), 
                  cells = "all")
testis_expression(LUSC_cluster3, cells = "all")
testis_expression(LUAD_cluster3, cells = "all")
testis_expression(LUAD_cluster4, cells = "all")
testis_expression(BRCA_cluster1, cells = "all")
testis_expression(BRCA_cluster2, cells = "all")
```

They mainly seem to, in a cluster, be expressed at the same time.

### MAGE genes

```{r MAGE_TCGA}
MAGE_genes <- filter(CT_genes, family == "MAGE")$external_gene_name

TCGA_expression(tumor = "all", 
                genes = MAGE_genes,
                units = "log_TPM")

CCLE_expression(genes = MAGE_genes,
                 type = c("lung", "skin", "bile_duct", "bladder", 
                          "colorectal", "lymphoma", "uterine",
                          "myeloma", "kidney", "pancreatic", "brain", 
                          "gastric", "breast", "bone", "head_and_neck", 
                          "ovarian", "sarcoma", "leukemia", "esophageal",
                          "neuroblastoma"), units = "log_TPM")

normal_tissues_mean_methylation(MAGE_genes, na.omit = TRUE)

```

## Single Cell analysis : timing of expression

In CTexploreR, we there is single cell data from the testis, we thus can 
analyse CT genes expression during spermatogenesis.

There is only data for 251 of our 298 CT genes.

NB : SSC, Spermatogonia and early spermatocytes are *premeiotic cells*. Late 
spermatocytes (between both meiosis), round spermatid, elongated spermatid and 
sperm are *postmeiotic cells*.

```{r repartition_sc}
genes_avail <- 
  CT_genes$external_gene_name[CT_genes$external_gene_name %in% unique(rownames(testis_sce))]
table(CT_genes$testis_cell_type)
table(CT_genes$testis_cell_type)/length(genes_avail)*100
```

41% of genes are mainly expressed pre-meioticly.

```{r sc_heatmap, message = FALSE}
germ_cells <- c("SSC", "Spermatogonia", "Early_spermatocyte",
                "Late_spermatocyte","Round_spermatid", "Elongated_spermatid",
                "Sperm1", "Sperm2")
somatic_cells <- c("Macrophage", "Endothelial", "Myoid", "Sertoli", "Leydig")

testis_sce_CT <- testis_sce[genes_avail, ]

mat <- SingleCellExperiment::logcounts(testis_sce_CT)

df_col <- data.frame(clusters = colData(testis_sce_CT)$clusters,
                     type = colData(testis_sce_CT)$type,
                     Donor = colData(testis_sce_CT)$Donor)
rownames(df_col) <- colnames(testis_sce_CT)

df_col <- df_col[order(df_col$type),]
df_col$lineage <- "Germ cells"
df_col$lineage[df_col$type %in% somatic_cells] <- "Somatic cells"

column_ha_type = HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  col = list(type = c("SSC" = "floralwhite", "Spermatogonia" = "moccasin",
                   "Early_spermatocyte" = "gold", 
                   "Late_spermatocyte" = "orange",
                   "Round_spermatid" = "red2", 
                   "Elongated_spermatid" = "darkred",
                   "Sperm1" = "violet", "Sperm2" = "purple", 
                   "Sertoli" = "gray", 
                   "Leydig" = "cadetblue2", "Myoid" = "springgreen3", 
                   "Macrophage" = "gray10",
                   "Endothelial" = "steelblue")),

  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

column_ha_lineage = HeatmapAnnotation(
  lineage = df_col$lineage,
  border = TRUE,
  col = list(lineage = c("Germ cells" = "salmon", "Somatic cells" = "cyan4")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

scale_lims <- c(0, quantile(rowMax(mat), 0.75))
top_annot <- c(column_ha_lineage, column_ha_type)

# Until here is what's in the function, hereunder is my addition/change in Heatmap()

CT_genes_avail <- filter(CT_genes, external_gene_name %in% genes_avail)

chr_mat <- as.matrix(CT_genes_avail$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes_avail$external_gene_name
row_ha_chr <- rowAnnotation(chr = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes_avail$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes_avail$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes_avail$regulated_by_methylation, CT_genes_avail$X_linked)

Heatmap(mat[genes_avail, rownames(df_col), drop = FALSE],
        name = "logCounts",
        column_title = "Expression in testis cells (scRNAseq)",
        column_split = df_col$type,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        show_column_names = FALSE,
        show_column_dend = FALSE,
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        col = colorRamp2(seq(scale_lims[1], scale_lims[2], length = 11),                 
                         legend_colors),
        top_annotation = top_annot,
        left_annotation = left_annot,
        heatmap_legend_param = legends_param)

```

We used these data to determine in which testis cell type
each gene is mostly expressed.

```{r main_cell_type_expression_barplot}
CT_genes$main_cell_type_expression <- factor(CT_genes$testis_cell_type,
                                                levels = c("SSC", 
                                               "Spermatogonia", 
                                               "Early_spermatocyte",
                                               "Late_spermatocyte",
                                               "Round_spermatid",
                                               "Elongated_spermatid",
                                               "Sperm1",
                                               "Sperm2",
                                               "Sertoli"))

CT_genes %>% 
  mutate(regulated_by_methylation = ifelse(regulated_by_methylation,
                                           "Regulated by methylation",
                                           "Not regulated by methylation")) %>%
  filter(!is.na(testis_cell_type)) %>%
  filter(main_cell_type_expression != "Sertoli") %>% 
  ggplot(aes(x = chr, fill = main_cell_type_expression,
             color = main_cell_type_expression)) +
  scale_fill_manual(values = c("lightyellow2", "moccasin", "gold", "orange","red2",
                               "darkred", "violet", "purple")) +
  scale_color_manual(values = c("lightyellow3", "navajowhite2", "gold", "orange","red2",
                               "darkred", "violet", "purple")) +
  geom_bar(stat = 'count') +
  facet_grid(main_cell_type_expression ~ regulated_by_methylation) +
  xlab("Chromosome") +
  ylab("Number of genes") +
  theme_bw() +
  theme(
    axis.text.x = element_text(color = "black", angle = 90, vjust = 0.5, size = 10),
    axis.title = element_text(size = 10, color = "gray10"),
    strip.text.y = element_blank())

```

14 genes are on the X chromosome but escape X chromosome inactivation and are
activates post-meioticly.

```{r escapees}
CT_genes %>% 
  filter(X_linked) %>% 
  filter(testis_cell_type %in% c("Late_spermatocyte","Round_spermatid", 
                                 "Elongated_spermatid", "Sperm1", "Sperm2"))

```

## Gene function

All CT genes function

```{r GO_ORA}
msigdbr(species = "Homo sapiens" , category = "C5") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

msigdbr(species = "Homo sapiens" , category = "H") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

go_ora <- enrichGO(gene = CT_genes$ensembl_gene_id,
                   keyType = "ENSEMBL",
                   OrgDb = org.Hs.eg.db,
                   ont = "all",
                   readable = TRUE)
as_tibble(go_ora)

as_tibble(go_ora) %>% 
  arrange(desc(Count)) %>% 
  head(12) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186)) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 4)

ora_to_plot <- as_tibble(simplify(go_ora))

# Combine the two cilium movement from BP
ora_to_plot[ora_to_plot$ID == "GO:0003341", "Count"] <- 
  ora_to_plot[ora_to_plot$ID == "GO:0003341", "Count"] +
  ora_to_plot[ora_to_plot$ID == "GO:0060294", "Count"]

ora_to_plot <- ora_to_plot %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  filter(ID != "GO:0060294") %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186))

ora_to_plot %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 3.7)
```

As we can see here, most of genes are indeed linked to functions from reproduction.
I represented here the 12 categories with the most genes, all from biological 
processes. However, they are enriched in 89 different GO terms.

Is there a difference between meth reg or not ?

```{r GO_ORA_meth}
go_ora_meth <- enrichGO(gene = 
                          filter(CT_genes, regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_meth <- simplify(go_ora_meth)

go_ora_not_meth <- enrichGO(gene = 
                          filter(CT_genes, !regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)

go_ora_not_meth <- simplify(go_ora_not_meth)
as_tibble(go_ora_not_meth)

go_ora_meth_plot <- as_tibble(go_ora_meth) %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/101,
                           ONTOLOGY == "MF"~ Count/107)) %>%  
  mutate(regulation = "Methylation")

go_ora_not_meth_plot <- as_tibble(go_ora_not_meth) %>% 
  arrange(desc(Count)) %>% 
  head(11) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/76,
                           ONTOLOGY == "CC"~ Count/93))%>%  
  mutate(regulation = "Not methylation")

rbind(go_ora_meth_plot, go_ora_not_meth_plot) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  facet_wrap(~ regulation) + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 6, color = "gray10"),
        axis.title.x = element_text(size = 10, color = "gray10"),
        axis.title.y = element_blank())+ 
  geom_text(aes(0, y = Description, label = ID),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 2)

```

We've also added a column tumor suppressor and oncogene in CTexploreR. 
These information come from Cancermine.

```{r C_TS}
table(CT_genes$oncogene) 
table(CT_genes$tumor_suppressor)

table(CT_genes$oncogene, CT_genes$tumor_suppressor) 
```

# SessionInfo

```{r}
sessionInfo()
```