### 2024-05-10 Paper_code-Database_CT.Rmd
Update and enhancement on database comparison

---
title: "Code used to generate paper results : Database comparison and CT characteristics"
author: "Julie Devis"
date: "2023-08-28"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

# Loading data and packages

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(Vennerable)
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(UpSetR)
library(ComplexHeatmap)
library(circlize)
library(SingleCellExperiment)
library(org.Hs.eg.db)
library(clusterProfiler)
library(msigdbr)
library(DOSE)
library(BiocParallel)
library(patchwork)
library(Biostrings)
```

Gene names/synonyms required for databases cleaning

```{r ensembl_synonym, warning=FALSE, message=FALSE}
ensembl <- biomaRt::useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
attributes_vector <- c("ensembl_gene_id", "external_gene_name",
                       "external_synonym", "gene_biotype",
                       "chromosome_name", "band", "start_position", "end_position",
                       "strand")
ensembl_gene_synonym <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))

ensembl_gene_synonym <- ensembl_gene_synonym %>%
  mutate(external_synonym = sub(pattern = "ORF", external_synonym, 
                                replacement = "orf"))

attributes_vector <- c("ensembl_gene_id", "external_gene_name")
ensembl_gene_names <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))

attributes_vector <- c("external_gene_name",
                       "external_synonym")
gene_synonym <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))
```

```{r load_data, warning=FALSE, message=FALSE}
# CCLE_data <- CTdata::CCLE_data()
# TCGA_TPM <- CTdata::TCGA_TPM()
# testis_sce <- CTdata::testis_sce()
####

load("../CTdata/eh_data/CCLE_data.rda")
load("../CTdata/eh_data/TCGA_TPM.rda")
load("../CTdata/eh_data/testis_sce.rda")
load("../CTdata/eh_data/mean_methylation_in_tissues.rda")

load("../CTdata/eh_data/all_genes.rda")
load("../CTdata/eh_data/HPA_cell_type_specificities.rda")
load("../CTdata/eh_data/CT_genes.rda")

CT_and_CTP_genes <- CT_genes
CT_genes <- (filter(CT_genes, CT_gene_type == "CT_gene"))
```

Common figures parameters

```{r fig_param}
legends_param <- list(
  labels_gp = gpar(col = "black", fontsize = 6),
  title_gp = gpar(col = "black", fontsize = 6),
  row_names_gp = gpar(fontsize = 4),
  annotation_name_side = "left")

legend_colors <- c("#5E4FA2", "#3288BD", "#66C2A5", "#ABDDA4", "#E6F598",
                   "#FFFFBF", "#FEE08B", "#FDAE61", "#F46D43", "#D53E4F",
                   "#9E0142")
chr_colors <- c("X-linked" = "deeppink", "Not X" = "royalblue1")

meth_colors <- c("Methylation" = "lightgreen", "Not methylation" = "gray")
```

# Pipeline figure values

Here is the code used to have all the values in the pipeline figure

```{r pipeline figures value}
table(all_genes$GTEX_category)
table(all_genes$multimapping_analysis)

table(all_genes$testis_specificity)
specific_to_pref <- all_genes %>%
  filter(GTEX_category == "testis_specific" |
           multimapping_analysis == "testis_specific") %>% 
  filter(testis_specificity == "testis_preferential")
table(specific_to_pref$not_detected_in_somatic_HPA)
table(specific_to_pref$CCLE_category)
table(specific_to_pref$TCGA_category)

all_genes %>% 
  filter(testis_specificity != "not_testis_specific") %>% 
  filter(CCLE_category == "activated" & 
           (TCGA_category == "multimapping_issue" | 
              TCGA_category == "activated")) %>% 
  pull(testis_specificity) %>% 
  table()

table(all_genes$CT_gene_type) #Before and after TSS verif
```

# Database comparison

## Lists cleaning

CT lists from other databases have been checked (using GTEx and our 
`GTEx_expression()` funtion and GeneCards) in order to remove duplicated
gene names or deprecated ones and allow comparison between databases.

### CTdatabase

Online list copied in a csv file, several lists exist so we combined them.

We checked gene names that were a concatenation of two genes (choice using
biomaRt synonyms to get the official one), checked which ones had the right names,
removed duplicated genes, verified lost genes and added back those that should 
be there.

```{r CTdatabase_cleaning, message = FALSE}
CTdatabase <- read_delim("data/CTdatabase1.csv", delim = ";", 
                         escape_double = FALSE, trim_ws = TRUE)
colnames(CTdatabase) <- c("Family", "Gene_Name", "Chromosomal_localization",
                          "CT_identifier")
CTdatabase_bis <- read_csv2("data/CTdatabase2.csv")
CTdatabase <- left_join(CTdatabase, CTdatabase_bis, 
                        by = c("Gene_Name" = "Gene_Symbol"))

CTdatabase_single <- CTdatabase %>%
  mutate(Gene_Name = sub(pattern = "/.*$", Gene_Name, replacement = ""))
CTdatabase_single <- CTdatabase_single %>%
  mutate(Gene_Name = sub(pattern = ",.*$", Gene_Name, replacement = ""))

CTdatabase_official_names <- 
  unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, 
                       external_gene_name)) %>%
  filter(external_gene_name %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA)
CTdatabase_synonym <- 
  ensembl_gene_synonym %>%
  filter(external_synonym %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_synonym) %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym)
CTdatabase_cleaned <- 
  rbind(CTdatabase_official_names, CTdatabase_synonym) %>% 
  left_join(CTdatabase_single)

duplicated_genes <- CTdatabase_cleaned$Gene_Name[duplicated(CTdatabase_cleaned$Gene_Name)]
bad_ids <- ensembl_gene_synonym %>%
  filter(external_gene_name %in% duplicated_genes | external_synonym %in% duplicated_genes) %>%
  filter(chromosome_name %in% grep(pattern = "H", x = chromosome_name, value = TRUE)) %>%
  pull(ensembl_gene_id)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  dplyr::filter(!ensembl_gene_id %in% bad_ids)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  filter(!ensembl_gene_id == "ENSG00000052126")
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!(ensembl_gene_id == "ENSG00000183305" & Gene_Name == "MAGEA2"))
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!ensembl_gene_id == "ENSG00000204648")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CSAG3B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CSAG2", "external_synonym"] <- "CSAG3B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT45A4")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CT45A3", "external_synonym"] <- "CT45A4"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "LAGE-1b")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAG2", "external_synonym"] <- "LAGE-1b"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT16.2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "PAGE5", "external_synonym"] <- "CT16.2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXB2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXB1", "external_synonym"] <- "SPANXB2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXE")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXD", "external_synonym"] <- "SPANXE"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1C")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1D")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1E")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE2B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "XAGE2", "external_synonym"] <- "XAGE2B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CTAGE-2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAGE1", "external_synonym"] <- "CTAGE-2"

CTdatabase_cleaned <- ensembl_gene_synonym %>%
  mutate(Gene_Name = external_synonym) %>%
  filter(external_synonym == "CXorf61") %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym) %>%
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "Cxorf61", 
                    c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name)) %>%
  filter(external_gene_name == "CCNA1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "cyclin A1", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "GOLGA6L2") %>%
  filter(ensembl_gene_id == "ENSG00000174450") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "GOLGAGL2 FA", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LYPD6B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC130576", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT62") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC196993", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT75") %>%
  filter(ensembl_gene_id == "ENSG00000291155") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC440934", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LINC01192") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC647107", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "TSPY1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC728137", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "SSX2B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "SSX2b", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
```

### Jamin's list

Excel file coming from supplemental data.

```{r Jamin_clean, message = FALSE}
Jamin_core_CT <- read_excel("data/Jamin_core_CT.xlsx")
Jamin_core_CT[Jamin_core_CT$Gene == "KIAA1211", "Gene"] <- "CRACD"
Jamin_core_CT[Jamin_core_CT$Gene == "CXorf67", "Gene"] <- "EZHIP"
```

### Wang's CTatlas

Excel file coming from supplemental data.

```{r Wang_cleaning, message = FALSE}
Wang_CT <- read_excel("data/Wang_Suppl_Data_3.xlsx", 
    sheet = "Supplementary Data 3B", skip = 1)
colnames(Wang_CT)[1] <- "ensembl_gene_id"

Wang_CT <- ensembl_gene_names %>% 
  filter(ensembl_gene_id %in% Wang_CT$ensembl_gene_id) %>%
  right_join(Wang_CT)

Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000181013", "external_gene_name"] <- "C17orf47"
Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000204293", "external_gene_name"] <- "OR8B2"
Wang_CT[Wang_CT$external_gene_name == "", "external_gene_name"] <- "RNASE11"
Wang_CT[Wang_CT$external_gene_name == "CHCT1", "external_gene_name"] <- "C17orf64"
Wang_CT[Wang_CT$external_gene_name == "PRSS40A", "external_gene_name"] <- "TISP43"
Wang_CT[Wang_CT$external_gene_name == "TEX56P", "external_gene_name"] <- "C6orf201"
Wang_CT[Wang_CT$external_gene_name == "SLC25A51P4", "external_gene_name"] <- "RP11-113D6.10"
Wang_CT[Wang_CT$external_gene_name == "TCP10L3", "external_gene_name"] <- "TCP10"
Wang_CT[Wang_CT$external_gene_name == "SCAND3", "external_gene_name"] <- "ZBED9"
```

### Carter's list

```{r Carter_clean, message = FALSE}
Carter_CT_list <- read_table("data/Carter_CT_list.txt", skip = 1)
Carter_CT <- Carter_CT_list[Carter_CT_list$CT_Expression,]

Carter_CT[Carter_CT$Gene == "ENSG00000261649", "Gene_Name"] <- "GOLGA6L7"
Carter_CT[Carter_CT$Gene == "ENSG00000239620", "Gene_Name"] <- "PRR20G"
Carter_CT[Carter_CT$Gene == "ENSG00000168148", "Gene_Name"] <- "H3-4"
Carter_CT[Carter_CT$Gene == "ENSG00000204296", "Gene_Name"] <- "TSBP1"
Carter_CT[Carter_CT$Gene == "ENSG00000180219", "Gene_Name"] <- "GARIN6"
Carter_CT[Carter_CT$Gene == "ENSG00000172717", "Gene_Name"] <- "GARIN2"
Carter_CT[Carter_CT$Gene == "ENSG00000174015", "Gene_Name"] <- "CBY2"
Carter_CT[Carter_CT$Gene == "ENSG00000224960", "Gene_Name"] <- "PPP4R3C"
```

### Burggeman's list

Excel file from supplemental data.

```{r Bruggeman_clean, message=FALSE}
Bruggeman_data <- read_excel("data/Bruggeman_suppl_data.xlsx", skip = 1,
                           sheet = "1D")

Bruggeman_official_names <- gene_synonym %>% 
  dplyr::select(external_gene_name) %>% 
  unique() %>% 
  filter(external_gene_name %in% Bruggeman_data$Gene) %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA)

Bruggeman_synonym <- gene_synonym %>%
  filter(external_synonym %in% Bruggeman_data$Gene) %>%
  mutate(Gene_Name = external_synonym) %>%
  dplyr::select(external_gene_name, Gene_Name, external_synonym)

Bruggeman_synonym <- Bruggeman_synonym[-which(Bruggeman_synonym$Gene_Name %in% 
                          Bruggeman_official_names$Gene_Name),]

Bruggeman_CT <- rbind(Bruggeman_official_names, Bruggeman_synonym)

lost <- Bruggeman_data[which(!Bruggeman_data$Gene %in% c(Bruggeman_CT$Gene_Name)), "Gene"]
colnames(lost) <- "external_gene_name"
lost$Gene_Name <- rep(NA, nrow(lost))
lost$external_synonym <- rep(NA, nrow(lost))

lost[lost$external_gene_name == "C21orf59", "Gene_Name"] <- "CFAP298"
lost[lost$external_gene_name == "C11orf57", "Gene_Name"] <- "NKAPD1"
lost[lost$external_gene_name == "C7orf55", "Gene_Name"] <- "FMC1"
lost[lost$external_gene_name == "C10orf12", "Gene_Name"] <- "LCOR"
lost[lost$external_gene_name == "RPL19P12", "Gene_Name"] <- "RPL19P12"
lost[lost$external_gene_name == "C16orf59", "Gene_Name"] <- "TEDC2"
lost[lost$external_gene_name == "TTTY15", "Gene_Name"] <- "USP9Y"
lost[lost$external_gene_name == "C17orf53", "Gene_Name"] <- "HROB"
lost[lost$external_gene_name == "C1orf112", "Gene_Name"] <- "FIRRM"
lost[lost$external_gene_name == "C12orf66", "Gene_Name"] <- "KICS2"
lost[lost$external_gene_name == "C9orf84", "Gene_Name"] <- "SHOC1"
lost[lost$external_gene_name == "C10orf25", "Gene_Name"] <- "ZNF22-AS1"
lost[lost$external_gene_name == "C20orf197", "Gene_Name"] <- "LINC02910"
lost[lost$external_gene_name == "C3orf67", "Gene_Name"] <- "CFAP20DC"
lost[lost$external_gene_name == "C8orf37", "Gene_Name"] <- "CFAP418"
lost[lost$external_gene_name == "C22orf34", "Gene_Name"] <- "MIR3667HG"

Bruggeman_CT <- rbind(Bruggeman_CT, lost)

missing_Bruggeman <- c("BMS1P4", "ADAM6", "ANXA2P3", "ARHGAP11B", "DPY19L2P2", 
                       "HLA-L", "PA2G4P4", "PIPSL", "PRKY", "YBX3P1", 
                       "RPL23AP53", "UQCRBP1", "RPL23P8", "MRS2P2", "PIN4P1", 
                       "SLC6A10P", "GUSBP2", "PPIEL", "LRRC37BP1", "MSL3P1", 
                       "PLEKHA8P1", "STAG3L1", "TCAM1P", "ZNF702P", "ZNF815P", 
                       "ATP6AP1L", "RPL21P44", "SEC14L1P1", "ZNF876P", 
                       "RPLP0P2", "FAM86JP", "FAM175A", "LACE1", "ATP5EP2", 
                       "WDR92", "TCTE3", "METTL20", "KIAA2022", "ZNRD1-AS1", 
                       "SGOL1", "FAM35DP", "MTL5", "TMEM14E", "MLLT4-AS1", 
                       "CCDC173", "KIAA1524", "WDR78", "LINC00476", "LYRM5", 
                       "HILS1", "CASC5", "KIAA1919", "CTAGE5", "FAM188B", 
                       "TMEM194B", "FAM122C", "PPP1R2P3", "KIAA0391", "SGOL2", 
                       "FAM19A3", "ZNF788", "RPL19P12", "FIRRM")

external_names_to_keep <- gene_synonym %>%
   filter(external_synonym %in% missing_Bruggeman) %>%
   filter(!external_gene_name %in% c("ATP5F1EP2", "POLR1HASP", "SHLD2P3", 
                                   "TMEM14EP", "H1-9P", "ZNF788P")) %>% 
   mutate(Gene_Name = external_gene_name)

Bruggeman_CT[Bruggeman_CT$external_synonym %in% 
                    external_names_to_keep$external_synonym, 
                  "Gene_Name"] <- external_names_to_keep$Gene_Name

Bruggeman_CT <- Bruggeman_CT %>% 
  dplyr::select(Gene_Name)
```

## CTexploreR data for selection pipeline

To characterise the differences between our database and other, we need the 
category we created in CTexploreR. For this, we have the object `all_genes` in 
CTdata that contains the CT analysis for all genes. More info in

Hereunder is what we used for our selection pipeline (coming from 
`make_all_genes_prelim.R` and `130_make_all_genes_and_CT_genes.R` in `CTdata`).

```{r all_genes, message = FALSE, warning = FALSE}
all_genes
```

From there, we filtered based on the testis_specificity ("testis_specific", which
is based on expression in health tissue and scRNA seq info from HPA), 
CCLE_category ("activated") and TCGA_category ("activated" or 
"multimapping_issue") to have our CT genes. Then, when wanting to validate TSS 
manually, we realised that for some genes, reads were not properly aligned to
exons which might reflect a poorly defined transcription in these regions and 
are hence likely unreliable.

Some genes were also characterized as Cancer-Testis preferential genes when 
testis specificity was less stringent

## CTexploreR VS CTdatabase

```{r CTdatabase_ours}
CTdatabase_ours <- Venn(list(CTdatabase = CTdatabase_cleaned$external_gene_name,
                             CTexploreR = CT_genes$external_gene_name))
gp <- VennThemes(compute.Venn(CTdatabase_ours))
gp[["Face"]][["11"]]$fill <-  "mistyrose"
gp[["Face"]][["01"]]$fill <-  "darkseagreen1"
gp[["Face"]][["10"]]$fill <-  "lightsteelblue1"
gp[["Set"]][["Set1"]]$col <-  "paleturquoise4"
gp[["Set"]][["Set2"]]$col <-  "darkseagreen4"
gp[["SetText"]][["Set1"]]$col <-  "paleturquoise4"
gp[["SetText"]][["Set2"]]$col <-  "darkseagreen4"
plot(CTdatabase_ours, gp = gp)

```

We find `r length(CTdatabase_ours@IntersectionSets[["11"]])/nrow(CTdatabase_cleaned)*100`
% of CTdatabase in CTexploreR, which is 
`r length(CTdatabase_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`% of our 
database.

Lost genes analysis

```{r CTdatabase_lost}
CTdatabase_lost <- all_genes %>%
  filter(external_gene_name %in% CTdatabase_ours@IntersectionSets[["10"]])

# 9 Genes are lost because not in any database

table(CTdatabase_lost$testis_specificity)

table(CTdatabase_lost$CT_gene_type)

table(CTdatabase_lost$not_detected_in_somatic_HPA)

table(CTdatabase_lost$TCGA_category)

table(CTdatabase_lost$CCLE_category)

table(CTdatabase_lost$TCGA_category, CTdatabase_lost$CCLE_category)

CTdatabase_lost_upset <- 
  list(`Not testis specific` = 
         filter(CTdatabase_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(CTdatabase_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(CTdatabase_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_CTdatabase <- fromList(CTdatabase_lost_upset)
upset(upset_CTdatabase)

 (nrow(CTdatabase_lost) - table(CTdatabase_lost$testis_specificity)[["testis_specific"]])/nrow(CTdatabase_lost)*100
```

`r (nrow(CTdatabase_lost) - table(CTdatabase_lost$testis_specificity)[["testis_specific"]])/nrow(CTdatabase_lost)*100`
% of these genes are not testis specific.

However `r table(CTdatabase_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(CTdatabase_lost) - table(CTdatabase_lost$TCGA_category, CTdatabase_lost$CCLE_category)["activated", "activated"])/nrow(CTdatabase_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

In their analysis, they had characterised gene specificity, some being not 
available, not found in testis, testis-restricted, testis-selective and 
testis/brain-restricted. Let's see how the lost genes qualify as they didn't
mention those were strictly testis specific.

```{r CTdatabase_classification}
CTdatabase_cleaned %>% 
  filter(external_gene_name %in% CTdatabase_lost$external_gene_name) %>% 
  pull(Classification) %>% 
  table()
```

We can see that most of them had no info or were testis-selective (I couldn't
find on website or paper how they selected categories).

## CTexploreR VS omics databases

```{r upset_plot}
core_ours <- Venn(list(Jamin = Jamin_core_CT$Gene, 
                       CTexploreR = CT_genes$external_gene_name))

Wang_ours <- Venn(list(Wang = Wang_CT$external_gene_name, 
                       CTexploreR = CT_genes$external_gene_name))

Carter_ours <- Venn(list(Carter_CT = Carter_CT$Gene_Name, 
                         CTexploreR = CT_genes$external_gene_name))

Bruggeman_ours <- Venn(list(Bruggeman = Bruggeman_CT$Gene_Name,
                            CTexploreR = CT_genes$external_gene_name))

gene_list <- list(CTexploreR = CT_genes$external_gene_name,
                  Carter = Carter_CT$Gene_Name,
                  Jamin = Jamin_core_CT$Gene, 
                  CTatlas = Wang_CT$external_gene_name,
                  Bruggeman = Bruggeman_CT$Gene_Name)

upset_omics <- fromList(gene_list)
upset(upset_omics)
```

4 in all, 60 in at least 3 databases

Lost genes analysis

```{r Jamin_lost}
plot(core_ours, gp = gp)

Jamin_lost <- all_genes %>%
  filter(external_gene_name %in% core_ours@IntersectionSets[["10"]])

table(Jamin_lost$testis_specificity)

table(Jamin_lost$CT_gene_type)

table(Jamin_lost$not_detected_in_somatic_HPA)

table(Jamin_lost$TCGA_category)

table(Jamin_lost$CCLE_category)

table(Jamin_lost$TCGA_category, Jamin_lost$CCLE_category)

Jamin_lost_upset <- 
  list(`Not testis specific` = 
         filter(Jamin_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Jamin_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Jamin_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Jamin <- fromList(Jamin_lost_upset)
upset(upset_Jamin)
```

We find `r length(core_ours@IntersectionSets[["11"]])/nrow(Jamin_core_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(core_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Jamin_lost) - table(Jamin_lost$testis_specificity)[["testis_specific"]])/nrow(Jamin_lost)*100`% of 
these genes are not testis specific.

However `r table(Jamin_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Jamin_lost) - table(Jamin_lost$TCGA_category, Jamin_lost$CCLE_category)["activated", "activated"])/nrow(Jamin_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Wang_lost}
plot(Wang_ours, gp = gp)

Wang_lost <- all_genes %>%
  filter(external_gene_name %in% Wang_ours@IntersectionSets[["10"]])

table(Wang_lost$testis_specificity)

table(Wang_lost$CT_gene_type)

table(Wang_lost$not_detected_in_somatic_HPA)

table(Wang_lost$TCGA_category)

table(Wang_lost$CCLE_category)

table(Wang_lost$TCGA_category, Wang_lost$CCLE_category)

Wang_lost_upset <- 
  list(`Not testis specific` = 
         filter(Wang_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Wang_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Wang_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Wang <- fromList(Wang_lost_upset)
upset(upset_Wang)
```

We find `r length(Wang_ours@IntersectionSets[["11"]])/nrow(Wang_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Wang_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Wang_lost) - table(Wang_lost$testis_specificity)[["testis_specific"]])/nrow(Wang_lost)*100`% of 
these genes are not testis specific.

However `r table(Wang_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Wang_lost) - table(Wang_lost$TCGA_category, Wang_lost$CCLE_category)["activated", "activated"])/nrow(Wang_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Carter_lost}
plot(Carter_ours, gp = gp)

Carter_lost <- all_genes %>%
  filter(external_gene_name %in% Carter_ours@IntersectionSets[["10"]])

table(Carter_lost$testis_specificity)

table(Carter_lost$CT_gene_type)

table(Carter_lost$not_detected_in_somatic_HPA)

table(Carter_lost$TCGA_category)

table(Carter_lost$CCLE_category)

table(Carter_lost$TCGA_category, Carter_lost$CCLE_category)

Carter_lost_upset <- 
  list(`Not testis specific` = 
         filter(Carter_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Carter_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Carter_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Carter <- fromList(Carter_lost_upset)
upset(upset_Carter)
```

We find `r length(Carter_ours@IntersectionSets[["11"]])/nrow(Carter_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Carter_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Carter_lost) - table(Carter_lost$testis_specificity)[["testis_specific"]])/nrow(Carter_lost)*100`% of 
these genes are not testis specific.

However `r table(Carter_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Carter_lost) - table(Carter_lost$TCGA_category, Carter_lost$CCLE_category)["activated", "activated"])/nrow(Carter_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Bruggeman_lost}
plot(Bruggeman_ours, gp = gp)

Bruggeman_lost <- all_genes %>%
  filter(external_gene_name %in% Bruggeman_ours@IntersectionSets[["10"]])

table(Bruggeman_lost$testis_specificity)

table(Bruggeman_lost$CT_gene_type)

table(Bruggeman_lost$not_detected_in_somatic_HPA)

table(Bruggeman_lost$TCGA_category)

table(Bruggeman_lost$CCLE_category)

table(Bruggeman_lost$TCGA_category, Bruggeman_lost$CCLE_category)

Bruggeman_lost_upset <- 
  list(`Not testis specific` = 
         filter(Bruggeman_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Bruggeman_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Bruggeman_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Bruggeman <- fromList(Bruggeman_lost_upset)
upset(upset_Bruggeman)
```

We find `r length(Bruggeman_ours@IntersectionSets[["11"]])/nrow(Bruggeman_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Bruggeman_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Bruggeman_lost) - table(Bruggeman_lost$testis_specificity)[["testis_specific"]])/nrow(Bruggeman_lost)*100`% of 
these genes are not testis specific.

However `r table(Bruggeman_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Bruggeman_lost) - table(Bruggeman_lost$TCGA_category, Bruggeman_lost$CCLE_category)["activated", "activated"])/nrow(Bruggeman_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

## Characterisation of differences with all databases

```{r all_databases}
common <- unique(c(core_ours@IntersectionSets[["11"]], 
                   CTdatabase_ours@IntersectionSets[["11"]], 
                   Wang_ours@IntersectionSets[["11"]], 
                   Carter_ours@IntersectionSets[["11"]],
                   Bruggeman_ours@IntersectionSets[["11"]]))

length(common)

length(common)/dim(CT_genes)[1] * 100

lost_list <- unique(c(core_ours@IntersectionSets[["10"]],
                      CTdatabase_ours@IntersectionSets[["10"]],
                      Wang_ours@IntersectionSets[["10"]],
                      Carter_ours@IntersectionSets[["10"]],
                      Bruggeman_ours@IntersectionSets[["10"]]))

lost <- all_genes %>%
  filter(external_gene_name %in% lost_list)

not_specific <- filter(lost, testis_specificity == "not_testis_specific")

GTEX_expression(not_specific$external_gene_name, units = "log_TPM")

somatic_testis <- filter(lost, not_detected_in_somatic_HPA == FALSE)

testis_expression(somatic_testis$external_gene_name, cells = "all")
HPA_cell_type_expression(somatic_testis$external_gene_name)

not_TCGA_activated <- filter(lost, TCGA_category != "activated" & 
                               TCGA_category != "multimapping_issue")

TCGA_expression(not_TCGA_activated$external_gene_name,
                tumor = "all",
                units = "log_TPM")

not_CCLE_activated <- filter(lost, CCLE_category  != "activated")

CCLE_expression(not_CCLE_activated$external_gene_name,
                  type = c("lung", "skin", "colorectal",
                           "gastric", "breast", "head_and_neck"),
                units = "log_TPM")

transcript_prob <- lost %>% 
  filter(testis_specificity == "testis_specific" |
           testis_specificity == "testis_preferential") %>%
  filter(TCGA_category == "activated" | TCGA_category == "multimapping_issue") %>% 
  filter(CCLE_category == "activated") %>% 
  dim()
```

`r length(common)` genes in our CTexploreR database are found in at least one of
the other database, which represents `r length(common)/dim(CT_genes)[1] * 100`%.

We have lost `r length(lost_list)` genes in total. Among them, 
`r dim(not_specific)[1]/length(lost_list)*100`% are not considered testis 
specific, `r dim(somatic_testis)[1]/length(lost_list)*100`% are expressed in 
somatic cells, `r dim(not_TCGA_activated)[1]/length(lost_list)*100`% 
are not activated in TCGA samples, 
`r dim(not_CCLE_activated)[1]/length(lost_list)*100`% are not activated in CCLE 
cell lines and `r transcript_prob[1]/length(lost_list)*100`% is lost due to 
transcripts problems.

What about new genes in CTexploreR

```{r new, warning = FALSE}
new <- CT_genes %>% 
  filter(!external_gene_name%in%common)

new

table(new$testis_specificity)
table(new$X_linked)
table(new$regulated_by_methylation)
table(new$X_linked, new$regulated_by_methylation)
TCGA_expression(tumor = "all", genes = new$external_gene_name, 
                units = "log_TPM")
TCGA_expression(tumor = "all", 
                genes = filter(new, X_linked & regulated_by_methylation)$external_gene_name, 
                units = "log_TPM")
```

There are `r dim(new)[1]` new CT genes in CTexploreR. These are all testis
specific and mainly on autosomes. Regulation by methylation is the majority of 
them.  There 
is only 11 new "major" CT that are on the X chromosome and regulated by 
methylation. CT45 are not that new.

Expression in tumours doesn't strike that much.

# CT genes characteristics

```{r basic}
table(CT_genes$testis_specificity)
table(CT_genes$transcript_biotype)
```

Most genes are testis specific 
(`r table(CT_genes$testis_specificity)[["testis_specific"]]/dim(CT_genes)[1]*100`%).
Most genes are mainly protein coding genes 
(`r table(CT_genes$transcript_biotype)["protein_coding"]/dim(CT_genes)[1]*100`%).

## X chromosome and regulated by methylation

In CTexploreR, genes have been characterised as regulated by methylation or not.

```{r X_meth}
table(CT_genes$X_linked)
table(CT_genes$regulated_by_methylation)
table(CT_genes$X_linked, CT_genes$testis_specificity)
table(CT_genes$X_linked, CT_genes$regulated_by_methylation)
```

Genes are enriched on the X chromosome 
(`r table(CT_genes$X_linked)[[2]]/dim(CT_genes)[1]*100`%). Also, 
`r table(CT_genes$regulated_by_methylation)[[2]]` genes have been flagged as 
regulated by methylation 
(`r table(CT_genes$regulated_by_methylation)[[2]]/dim(CT_genes)[1]*100`%). 
It is interesting to study the link between these two characteristics.

On the chromosome X, there is a clear enrichment of CT genes regulated by 
methylation (`r table(CT_genes$X_linked, CT_genes$regulated_by_methylation)[2, 2]`/`r table(CT_genes$X_linked)[[2]]` 
chrX genes or `r table(CT_genes$X_linked, CT_genes$regulated_by_methylation)[2, 2]`/`r table(CT_genes$regulated_by_methylation)[[2]]` 
genes regulated by methylation).

Let's check that with a statistical test, I here want to see if, on there is an 
enrichment of genes regulated by methylation on the X chromosome. I need to do a
Pearson Chi square test (to know if observed proportion differ from expected 
proportion). It is a statistical method to determine if two categorical variables
have a significant correlation between them. I can directly put a matrix 
(my table) in the function

```{r chisq}
chisq.test(table(CT_genes$X_linked, CT_genes$regulated_by_methylation))
```

```{r X_meth_fig}
CT_genes$chr_factor <- factor(CT_genes$chr,
                       levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9",
                                  "10", "11", "12", "13", "14", "15", "16", "17", 
                                  "18", "19", "20", "21", "22", "X", "Y"))

CT_genes %>% 
  mutate(regulated_by_methylation = ifelse(regulated_by_methylation,
                                           "Regulated by methylation",
                                           "Not regulated by methylation")) %>% 
  ggplot(aes(x = chr_factor, fill = X_linked)) +
  geom_bar(stat = 'count') +
  scale_fill_manual(values = c("royalblue1", "deeppink")) +
  facet_wrap(~ regulated_by_methylation, ncol = 2) +
  theme_bw() +
  xlab("Chromosome") +
  ylab("Number of genes") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "none",
        axis.title = element_text(size = 10, color = "gray10"))
```

There is indeed a significative link between regulation by methylation and being
on the X chromosome. There is thus an enrichment of CT genes regulated by 
methylation on the X chromosome (and inversely).

## Tumour and methylation

Using CTexploreR functions, we can explore all CT genes or focus on some 
potential targets.

### All CT

For these heatmaps, the code comes from the function but has been copies to add 
some annotations.

```{r common_annot}
chr_mat <- as.matrix(CT_genes$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes$external_gene_name
row_ha_chr <- rowAnnotation(chr_factor = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr_factor = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes$regulated_by_methylation, CT_genes$X_linked)
```

```{r TCGA_heatmap, message = FALSE}
database <- TCGA_TPM
database$tumor <- sub(pattern = 'TCGA-', x = database$project_id, '')
database$type <- "Tumor"
database$type[database$shortLetterCode == "NT"] <- "Peritumoral"
database <- database[, order(database$tumor, database$type)]

genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]
database <- database[, database$type == "Tumor"]

column_ha_tumor <- HeatmapAnnotation(
  Tumor = database$tumor,
  border = TRUE,
  col = list(Tumor = c("BRCA" = "midnightblue", "COAD" = "darkorchid2",
                 "ESCA" = "gold", "HNSC" = "deeppink2",
                 "LUAD" = "seagreen", "LUSC" = "seagreen3",
                 "SKCM" = "red3")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

split_by <- factor(database$tumor)
col_annot <- column_ha_tumor

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

Heatmap(mat[, , drop = FALSE],
        name = "logTPM",
        column_title = paste0("Expression in TCGA samples (all)"),
        column_split = split_by,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                         legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = col_annot,
        left_annotation = left_annot)

```

```{r CCLE_heatmap}
database <- CCLE_data
database$type <- tolower(database$type)
genes <- CT_genes$external_gene_name
database <- database[rowData(database)$external_gene_name %in% genes, ]
database <- database[match(genes, rowData(database)$external_gene_name), ]

mat <- log1p(assay(database))
rownames(mat) <- rowData(database)$external_gene_name

df_col <- data.frame("cell_line" = colData(database)$cell_line_name,
                     "type" = as.factor(colData(database)$type))
rownames(df_col) <- rownames(colData(database))
df_col <- df_col[order(df_col$type), ]

column_ha_type <- HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param,
  col = list(type = c("lung" = "seagreen3", "skin" = "red3",
                 "bile_duct" = "mediumpurple1", "bladder" = "mistyrose2",
                 "colorectal" = "plum", "lymphoma" = "steelblue1",
                 "uterine" = "darkorange4", "myeloma" = "turquoise3", 
                 "kidney" = "thistle4",
                 "pancreatic" = "darkmagenta", "brain" = "palegreen2",
                 "gastric" = "wheat3", "breast" = "midnightblue",
                 "bone" = "sienna1", "head_and_neck" = "deeppink2",
                 "ovarian" = "tan3", "sarcoma" = "lightcoral",
                 "leukemia" = "steelblue4", "esophageal"= "khaki",
                 "neuroblastoma" = "olivedrab1")))

Heatmap(mat[, rownames(df_col), drop = FALSE],
        name = "logTPM",
        column_title = "Gene Expression in tumor cell lines (CCLE)",
        column_split = factor(df_col$type),
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        col = colorRamp2(seq(0, max(mat), length = 11),
                          legend_colors),
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        show_row_dend = FALSE,
        show_column_names = FALSE,
        cluster_columns = TRUE,
        show_column_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        heatmap_legend_param = legends_param,
        top_annotation = c(column_ha_type),
        left_annotation = left_annot)
```

Next graph was removed from paper

```{r all_meth}
genes <- CT_genes$external_gene_name
database <- mean_methylation_in_tissues[rownames(mean_methylation_in_tissues) %in% genes]

mat <- na.omit(assay(database))
clustering_option <- TRUE

row_ha_chr_meth <- rowAnnotation(chr = chr_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

row_ha_reg_meth <- rowAnnotation(regulation = regulation_mat[rownames(mat),],
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot_meth <- c(row_ha_chr_meth, row_ha_reg_meth, gap = unit(1, "mm"))

split_meth <- data.frame(filter(CT_genes, external_gene_name %in% rownames(mat))$regulated_by_methylation, 
                    filter(CT_genes, external_gene_name %in% rownames(mat))$X_linked)

Heatmap(mat,
        column_title = 'Promoter mean methylation level by tissue',
        name = 'Meth',
        col = colorRamp2(c(1:100),
                         colorRampPalette(c("moccasin","dodgerblue4"))(100)),
        na_col = "gray80",
        cluster_rows = clustering_option,
        cluster_columns = FALSE,
        row_split = split_meth,
        row_title_gp = gpar(fontsize = 0),
        show_row_names = TRUE,
        show_heatmap_legend = TRUE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 3),
        column_names_gp = gpar(fontsize = 8),
        column_names_side = "bottom",
        row_names_side = "right",
        left_annotation = left_annot_meth)

```

### MAGE genes

```{r MAGE_TCGA}
MAGE_genes <- filter(CT_genes, family == "MAGE")$external_gene_name

TCGA_expression(tumor = "all", 
                genes = MAGE_genes,
                units = "log_TPM")

CCLE_expression(genes = MAGE_genes,
                 type = c("lung", "skin", "bile_duct", "bladder", 
                          "colorectal", "lymphoma", "uterine",
                          "myeloma", "kidney", "pancreatic", "brain", 
                          "gastric", "breast", "bone", "head_and_neck", 
                          "ovarian", "sarcoma", "leukemia", "esophageal",
                          "neuroblastoma"), units = "log_TPM")

normal_tissues_mean_methylation(MAGE_genes, na.omit = TRUE)

```

## Single Cell analysis : timing of expression

In CTexploreR, we there is single cell data from the testis, we thus can 
analyse CT genes expression during spermatogenesis.

There is only data for `r sum(CT_genes$external_gene_name %in% unique(rownames(testis_sce)))` of our `r nrow(CT_genes)` CT genes.

NB : SSC, Spermatogonia and early spermatocytes are *premeiotic cells*. Late 
spermatocytes (between both meiosis), round spermatid, elongated spermatid and 
sperm are *postmeiotic cells*.

```{r repartition_sc}
genes_avail <- 
  CT_genes$external_gene_name[CT_genes$external_gene_name %in% unique(rownames(testis_sce))]
table(CT_genes$testis_cell_type)
table(CT_genes$testis_cell_type)/length(genes_avail)*100
```

`r sum(table(CT_genes$testis_cell_type)[c("Early_spermatocyte", "Spermatogonia", "SSC")])/length(genes_avail)*100`
% of genes are mainly expressed pre-meioticly.

```{r sc_heatmap, message = FALSE}
germ_cells <- c("SSC", "Spermatogonia", "Early_spermatocyte",
                "Late_spermatocyte","Round_spermatid", "Elongated_spermatid",
                "Sperm1", "Sperm2")
somatic_cells <- c("Macrophage", "Endothelial", "Myoid", "Sertoli", "Leydig")

testis_sce_CT <- testis_sce[genes_avail, ]

mat <- SingleCellExperiment::logcounts(testis_sce_CT)

df_col <- data.frame(clusters = colData(testis_sce_CT)$clusters,
                     type = colData(testis_sce_CT)$type,
                     Donor = colData(testis_sce_CT)$Donor)
rownames(df_col) <- colnames(testis_sce_CT)

df_col <- df_col[order(df_col$type),]
df_col$lineage <- "Germ cells"
df_col$lineage[df_col$type %in% somatic_cells] <- "Somatic cells"

column_ha_type = HeatmapAnnotation(
  type = df_col$type,
  border = TRUE,
  col = list(type = c("SSC" = "floralwhite", "Spermatogonia" = "moccasin",
                   "Early_spermatocyte" = "gold", 
                   "Late_spermatocyte" = "orange",
                   "Round_spermatid" = "red2", 
                   "Elongated_spermatid" = "darkred",
                   "Sperm1" = "violet", "Sperm2" = "purple", 
                   "Sertoli" = "gray", 
                   "Leydig" = "cadetblue2", "Myoid" = "springgreen3", 
                   "Macrophage" = "gray10",
                   "Endothelial" = "steelblue")),

  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

column_ha_lineage = HeatmapAnnotation(
  lineage = df_col$lineage,
  border = TRUE,
  col = list(lineage = c("Germ cells" = "salmon", "Somatic cells" = "cyan4")),
  annotation_name_gp = gpar(fontsize = 8),
  annotation_legend_param = legends_param)

scale_lims <- c(0, quantile(rowMax(mat), 0.75))
top_annot <- c(column_ha_lineage, column_ha_type)

# Until here is what's in the function, hereunder is my addition/change in Heatmap()

CT_genes_avail <- filter(CT_genes, external_gene_name %in% genes_avail)

chr_mat <- as.matrix(CT_genes_avail$X_linked)
chr_mat <- ifelse(chr_mat == TRUE, "X-linked", "Not X")
rownames(chr_mat) <- CT_genes_avail$external_gene_name
row_ha_chr <- rowAnnotation(chr = chr_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(chr = chr_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

regulation_mat <- as.matrix(CT_genes_avail$regulated_by_methylation)
regulation_mat <- ifelse(regulation_mat == TRUE, "Methylation",
                         "Not methylation")
rownames(regulation_mat) <- CT_genes_avail$external_gene_name

row_ha_reg <- rowAnnotation(regulation = regulation_mat,
                            annotation_legend_param = legends_param,
                            simple_anno_size = unit(0.5, "cm"),
                            col = list(regulation = meth_colors),
                            annotation_name_gp = gpar(fontsize = 8),
                            annotation_name_side = "top")

left_annot <- c(row_ha_chr, row_ha_reg, gap = unit(1, "mm"))

split <- data.frame(CT_genes_avail$regulated_by_methylation, CT_genes_avail$X_linked)

Heatmap(mat[genes_avail, rownames(df_col), drop = FALSE],
        name = "logCounts",
        column_title = "Expression in testis cells (scRNAseq)",
        column_split = df_col$type,
        row_split = split,
        row_title_gp = gpar(fontsize = 0),
        show_column_names = FALSE,
        show_column_dend = FALSE,
        clustering_method_rows = "ward.D",
        clustering_method_columns = "ward.D",
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        show_row_dend = FALSE,
        row_names_gp = gpar(fontsize = 4),
        col = colorRamp2(seq(scale_lims[1], scale_lims[2], length = 11),                 
                         legend_colors),
        top_annotation = top_annot,
        left_annotation = left_annot,
        heatmap_legend_param = legends_param)

```

## Gene function

All CT genes function

```{r GO_ORA}
msigdbr(species = "Homo sapiens" , category = "C5") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

msigdbr(species = "Homo sapiens" , category = "H") %>% 
  filter(gene_symbol %in% CT_genes$external_gene_name) %>% 
  pull(gene_symbol) %>% 
  unique() %>% 
  length()

go_ora <- enrichGO(gene = CT_genes$ensembl_gene_id,
                   keyType = "ENSEMBL",
                   OrgDb = org.Hs.eg.db,
                   ont = "all",
                   readable = TRUE)
as_tibble(go_ora)

as_tibble(go_ora) %>% 
  arrange(desc(Count)) %>% 
  head(12) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/177,
                           ONTOLOGY == "CC"~ Count/197,
                           ONTOLOGY == "MF"~ Count/186)) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 4)

ora_to_plot <- as_tibble(simplify(go_ora))

ora_to_plot <- ora_to_plot %>% 
  arrange(desc(Count)) %>% 
  head(9) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/102,
                           ONTOLOGY == "MF"~ Count/186))

ora_to_plot %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  theme(axis.text.y = element_blank(),
        legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title = element_text(size = 10, color = "gray10"))+ 
  geom_text(aes(0, y = Description, label = Description),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 3.7)
```

As we can see here, most of genes are indeed linked to functions from reproduction.
I represented here the 12 categories with the most genes, all from biological 
processes. However, they are enriched in `r nrow(as_tibble(go_ora))` different GO terms.

Is there a difference between meth reg or not ?

```{r GO_ORA_meth}
go_ora_meth <- enrichGO(gene = 
                          filter(CT_genes, regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_meth <- simplify(go_ora_meth)
as_tibble(go_ora_meth)

go_ora_not_meth <- enrichGO(gene = 
                          filter(CT_genes, !regulated_by_methylation)$ensembl_gene_id,
                        keyType = "ENSEMBL",
                        OrgDb = org.Hs.eg.db,
                        ont = "all",
                        readable = TRUE)
go_ora_not_meth <- simplify(go_ora_not_meth)
as_tibble(go_ora_not_meth)

go_ora_meth_plot <- as_tibble(go_ora_meth) %>% 
  arrange(desc(Count)) %>% 
  head(10) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/73,
                           ONTOLOGY == "MF"~ Count/80,
                           ONTOLOGY == "CC"~ Count/78)) %>%  
  mutate(regulation = "Methylation")

go_ora_not_meth_plot <- as_tibble(go_ora_not_meth) %>% 
  arrange(desc(Count)) %>% 
  head(8) %>% 
  mutate(Ratio = case_when(ONTOLOGY == "BP"~ Count/29,
                           ONTOLOGY == "MF"~ Count/30))%>%  
  mutate(regulation = "Not methylation")

rbind(go_ora_meth_plot, go_ora_not_meth_plot) %>% 
  ggplot(aes(x = Ratio, y = Description, fill = Description)) +
  geom_col() +
  theme_bw() +
  ylab("GO term") +
  xlab("Gene Ratio") +
  facet_wrap(~ regulation) + 
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.text.y = element_text(size = 6, color = "gray10"),
        axis.title.x = element_text(size = 10, color = "gray10"),
        axis.title.y = element_blank())+ 
  geom_text(aes(0, y = Description, label = ID),
            hjust = 0,
            nudge_x = 0.005,
            colour = "floralwhite",
            size = 2)

```

We've also added a column tumor suppressor and oncogene in CTexploreR. 
These information come from Cancermine.

```{r C_TS}
table(CT_genes$oncogene) 
table(CT_genes$tumor_suppressor)

table(CT_genes$oncogene, CT_genes$tumor_suppressor) 
```

# SessionInfo

```{r}
sessionInfo()
```