### 2024-05-14 Paper_code-Database_Comparison.Rmd

---
title: "Code used to generate paper results : Database comparison and CT characteristics"
author: "Julie Devis"
date: "2023-08-28"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

# Loading data and packages

```{r library, message = FALSE}
library(readxl)
library(readr)
library(CTexploreR)
library(Vennerable)
library(biomaRt)
library(tidyverse)
library(SummarizedExperiment)
library(UpSetR)
```

Gene names/synonyms required for databases cleaning

```{r ensembl_synonym, warning=FALSE, message=FALSE}
ensembl <- biomaRt::useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
attributes_vector <- c("ensembl_gene_id", "external_gene_name",
                       "external_synonym", "gene_biotype",
                       "chromosome_name", "band", "start_position", "end_position",
                       "strand")
ensembl_gene_synonym <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))

ensembl_gene_synonym <- ensembl_gene_synonym %>%
  mutate(external_synonym = sub(pattern = "ORF", external_synonym, 
                                replacement = "orf"))

attributes_vector <- c("ensembl_gene_id", "external_gene_name")
ensembl_gene_names <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))

attributes_vector <- c("external_gene_name",
                       "external_synonym")
gene_synonym <- as_tibble(getBM(attributes = attributes_vector, mart = ensembl))
```

```{r load_data, warning=FALSE, message=FALSE}
load("../CTdata/eh_data/all_genes.rda")
load("../CTdata/eh_data/CT_genes.rda")

CT_and_CTP_genes <- CT_genes
CT_genes <- (filter(CT_genes, CT_gene_type == "CT_gene"))

upset_text_size <- c(1.5, 1.5, 1.5, 1.5, 1.8, 2)
 #c(intersection size title, intersection size tick labels, set size title, 
 # set size tick labels, set names, numbers above bars)
```

# Database comparison

## Lists cleaning

CT lists from other databases have been checked (using GTEx and our 
`GTEx_expression()` funtion and GeneCards) in order to remove duplicated
gene names or deprecated ones and allow comparison between databases.

### CTdatabase

Online list copied in a csv file, several lists exist so we combined them.

We checked gene names that were a concatenation of two genes (choice using
biomaRt synonyms to get the official one), checked which ones had the right names,
removed duplicated genes, verified lost genes and added back those that should 
be there.

```{r CTdatabase_cleaning, message = FALSE}
CTdatabase <- read_delim("data/CTdatabase1.csv", delim = ";", 
                         escape_double = FALSE, trim_ws = TRUE)
colnames(CTdatabase) <- c("Family", "Gene_Name", "Chromosomal_localization",
                          "CT_identifier")
CTdatabase_bis <- read_csv2("data/CTdatabase2.csv")
CTdatabase <- left_join(CTdatabase, CTdatabase_bis, 
                        by = c("Gene_Name" = "Gene_Symbol"))

CTdatabase_single <- CTdatabase %>%
  mutate(Gene_Name = sub(pattern = "/.*$", Gene_Name, replacement = ""))
CTdatabase_single <- CTdatabase_single %>%
  mutate(Gene_Name = sub(pattern = ",.*$", Gene_Name, replacement = ""))

CTdatabase_official_names <- 
  unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, 
                       external_gene_name)) %>%
  filter(external_gene_name %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA)
CTdatabase_synonym <- 
  ensembl_gene_synonym %>%
  filter(external_synonym %in% CTdatabase_single$Gene_Name) %>%
  mutate(Gene_Name = external_synonym) %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym)
CTdatabase_cleaned <- 
  rbind(CTdatabase_official_names, CTdatabase_synonym) %>% 
  left_join(CTdatabase_single)

duplicated_genes <- CTdatabase_cleaned$Gene_Name[duplicated(CTdatabase_cleaned$Gene_Name)]
bad_ids <- ensembl_gene_synonym %>%
  filter(external_gene_name %in% duplicated_genes | external_synonym %in% duplicated_genes) %>%
  filter(chromosome_name %in% grep(pattern = "H", x = chromosome_name, value = TRUE)) %>%
  pull(ensembl_gene_id)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  dplyr::filter(!ensembl_gene_id %in% bad_ids)
CTdatabase_cleaned <- CTdatabase_cleaned %>%
  filter(!ensembl_gene_id == "ENSG00000052126")
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!(ensembl_gene_id == "ENSG00000183305" & Gene_Name == "MAGEA2"))
CTdatabase_cleaned <- CTdatabase_cleaned %>% 
  filter(!ensembl_gene_id == "ENSG00000204648")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CSAG3B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CSAG2", "external_synonym"] <- "CSAG3B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT45A4")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CT45A3", "external_synonym"] <- "CT45A4"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "LAGE-1b")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAG2", "external_synonym"] <- "LAGE-1b"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CT16.2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "PAGE5", "external_synonym"] <- "CT16.2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXB2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXB1", "external_synonym"] <- "SPANXB2"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "SPANXE")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "SPANXD", "external_synonym"] <- "SPANXE"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1C")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1D")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE1E")
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "XAGE2B")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "XAGE2", "external_synonym"] <- "XAGE2B"
CTdatabase_cleaned <- filter(CTdatabase_cleaned, Gene_Name != "CTAGE-2")
CTdatabase_cleaned[CTdatabase_cleaned$Gene_Name == "CTAGE1", "external_synonym"] <- "CTAGE-2"

CTdatabase_cleaned <- ensembl_gene_synonym %>%
  mutate(Gene_Name = external_synonym) %>%
  filter(external_synonym == "CXorf61") %>%
  dplyr::select(ensembl_gene_id, external_gene_name, Gene_Name, external_synonym) %>%
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "Cxorf61", 
                    c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name)) %>%
  filter(external_gene_name == "CCNA1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "cyclin A1", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "GOLGA6L2") %>%
  filter(ensembl_gene_id == "ENSG00000174450") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "GOLGAGL2 FA", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LYPD6B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC130576", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT62") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC196993", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "CT75") %>%
  filter(ensembl_gene_id == "ENSG00000291155") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC440934", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "LINC01192") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC647107", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "TSPY1") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "LOC728137", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
CTdatabase_cleaned <- unique(dplyr::select(ensembl_gene_synonym, ensembl_gene_id, external_gene_name))%>%
  filter(external_gene_name == "SSX2B") %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA) %>% 
  cbind(CTdatabase_single[CTdatabase_single$Gene_Name == "SSX2b", 
                          c("Family", "Chromosomal_localization", "CT_identifier", "Classification")]) %>% 
  rbind(CTdatabase_cleaned)
```

### Jamin's list

Excel file coming from supplemental data.

```{r Jamin_clean, message = FALSE}
Jamin_core_CT <- read_excel("data/Jamin_core_CT.xlsx")
Jamin_core_CT[Jamin_core_CT$Gene == "KIAA1211", "Gene"] <- "CRACD"
Jamin_core_CT[Jamin_core_CT$Gene == "CXorf67", "Gene"] <- "EZHIP"
```

### Wang's CTatlas

Excel file coming from supplemental data.

```{r Wang_cleaning, message = FALSE}
Wang_CT <- read_excel("data/Wang_Suppl_Data_3.xlsx", 
    sheet = "Supplementary Data 3B", skip = 1)
colnames(Wang_CT)[1] <- "ensembl_gene_id"

Wang_CT <- ensembl_gene_names %>% 
  filter(ensembl_gene_id %in% Wang_CT$ensembl_gene_id) %>%
  right_join(Wang_CT)

Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000181013", "external_gene_name"] <- "C17orf47"
Wang_CT[Wang_CT$ensembl_gene_id == "ENSG00000204293", "external_gene_name"] <- "OR8B2"
Wang_CT[Wang_CT$external_gene_name == "", "external_gene_name"] <- "RNASE11"
Wang_CT[Wang_CT$external_gene_name == "CHCT1", "external_gene_name"] <- "C17orf64"
Wang_CT[Wang_CT$external_gene_name == "PRSS40A", "external_gene_name"] <- "TISP43"
Wang_CT[Wang_CT$external_gene_name == "TEX56P", "external_gene_name"] <- "C6orf201"
Wang_CT[Wang_CT$external_gene_name == "SLC25A51P4", "external_gene_name"] <- "RP11-113D6.10"
Wang_CT[Wang_CT$external_gene_name == "TCP10L3", "external_gene_name"] <- "TCP10"
Wang_CT[Wang_CT$external_gene_name == "SCAND3", "external_gene_name"] <- "ZBED9"
```

### Carter's list

```{r Carter_clean, message = FALSE}
Carter_CT_list <- read_table("data/Carter_CT_list.txt", skip = 1)
Carter_CT <- Carter_CT_list[Carter_CT_list$CT_Expression,]

Carter_CT[Carter_CT$Gene == "ENSG00000261649", "Gene_Name"] <- "GOLGA6L7"
Carter_CT[Carter_CT$Gene == "ENSG00000239620", "Gene_Name"] <- "PRR20G"
Carter_CT[Carter_CT$Gene == "ENSG00000168148", "Gene_Name"] <- "H3-4"
Carter_CT[Carter_CT$Gene == "ENSG00000204296", "Gene_Name"] <- "TSBP1"
Carter_CT[Carter_CT$Gene == "ENSG00000180219", "Gene_Name"] <- "GARIN6"
Carter_CT[Carter_CT$Gene == "ENSG00000172717", "Gene_Name"] <- "GARIN2"
Carter_CT[Carter_CT$Gene == "ENSG00000174015", "Gene_Name"] <- "CBY2"
Carter_CT[Carter_CT$Gene == "ENSG00000224960", "Gene_Name"] <- "PPP4R3C"
```

### Burggeman's list

Excel file from supplemental data.

```{r Bruggeman_clean, message=FALSE}
Bruggeman_data <- read_excel("data/Bruggeman_suppl_data.xlsx", skip = 1,
                           sheet = "1D")

Bruggeman_official_names <- gene_synonym %>% 
  dplyr::select(external_gene_name) %>% 
  unique() %>% 
  filter(external_gene_name %in% Bruggeman_data$Gene) %>%
  mutate(Gene_Name = external_gene_name) %>%
  mutate(external_synonym = NA)

Bruggeman_synonym <- gene_synonym %>%
  filter(external_synonym %in% Bruggeman_data$Gene) %>%
  mutate(Gene_Name = external_synonym) %>%
  dplyr::select(external_gene_name, Gene_Name, external_synonym)

Bruggeman_synonym <- Bruggeman_synonym[-which(Bruggeman_synonym$Gene_Name %in% 
                          Bruggeman_official_names$Gene_Name),]

Bruggeman_CT <- rbind(Bruggeman_official_names, Bruggeman_synonym)

lost <- Bruggeman_data[which(!Bruggeman_data$Gene %in% c(Bruggeman_CT$Gene_Name)), "Gene"]
colnames(lost) <- "external_gene_name"
lost$Gene_Name <- rep(NA, nrow(lost))
lost$external_synonym <- rep(NA, nrow(lost))

lost[lost$external_gene_name == "C21orf59", "Gene_Name"] <- "CFAP298"
lost[lost$external_gene_name == "C11orf57", "Gene_Name"] <- "NKAPD1"
lost[lost$external_gene_name == "C7orf55", "Gene_Name"] <- "FMC1"
lost[lost$external_gene_name == "C10orf12", "Gene_Name"] <- "LCOR"
lost[lost$external_gene_name == "RPL19P12", "Gene_Name"] <- "RPL19P12"
lost[lost$external_gene_name == "C16orf59", "Gene_Name"] <- "TEDC2"
lost[lost$external_gene_name == "TTTY15", "Gene_Name"] <- "USP9Y"
lost[lost$external_gene_name == "C17orf53", "Gene_Name"] <- "HROB"
lost[lost$external_gene_name == "C1orf112", "Gene_Name"] <- "FIRRM"
lost[lost$external_gene_name == "C12orf66", "Gene_Name"] <- "KICS2"
lost[lost$external_gene_name == "C9orf84", "Gene_Name"] <- "SHOC1"
lost[lost$external_gene_name == "C10orf25", "Gene_Name"] <- "ZNF22-AS1"
lost[lost$external_gene_name == "C20orf197", "Gene_Name"] <- "LINC02910"
lost[lost$external_gene_name == "C3orf67", "Gene_Name"] <- "CFAP20DC"
lost[lost$external_gene_name == "C8orf37", "Gene_Name"] <- "CFAP418"
lost[lost$external_gene_name == "C22orf34", "Gene_Name"] <- "MIR3667HG"

Bruggeman_CT <- rbind(Bruggeman_CT, lost)

missing_Bruggeman <- c("BMS1P4", "ADAM6", "ANXA2P3", "ARHGAP11B", "DPY19L2P2", 
                       "HLA-L", "PA2G4P4", "PIPSL", "PRKY", "YBX3P1", 
                       "RPL23AP53", "UQCRBP1", "RPL23P8", "MRS2P2", "PIN4P1", 
                       "SLC6A10P", "GUSBP2", "PPIEL", "LRRC37BP1", "MSL3P1", 
                       "PLEKHA8P1", "STAG3L1", "TCAM1P", "ZNF702P", "ZNF815P", 
                       "ATP6AP1L", "RPL21P44", "SEC14L1P1", "ZNF876P", 
                       "RPLP0P2", "FAM86JP", "FAM175A", "LACE1", "ATP5EP2", 
                       "WDR92", "TCTE3", "METTL20", "KIAA2022", "ZNRD1-AS1", 
                       "SGOL1", "FAM35DP", "MTL5", "TMEM14E", "MLLT4-AS1", 
                       "CCDC173", "KIAA1524", "WDR78", "LINC00476", "LYRM5", 
                       "HILS1", "CASC5", "KIAA1919", "CTAGE5", "FAM188B", 
                       "TMEM194B", "FAM122C", "PPP1R2P3", "KIAA0391", "SGOL2", 
                       "FAM19A3", "ZNF788", "RPL19P12", "FIRRM")

external_names_to_keep <- gene_synonym %>%
   filter(external_synonym %in% missing_Bruggeman) %>%
   filter(!external_gene_name %in% c("ATP5F1EP2", "POLR1HASP", "SHLD2P3", 
                                   "TMEM14EP", "H1-9P", "ZNF788P")) %>% 
   mutate(Gene_Name = external_gene_name)

Bruggeman_CT[Bruggeman_CT$external_synonym %in% 
                    external_names_to_keep$external_synonym, 
                  "Gene_Name"] <- external_names_to_keep$Gene_Name

Bruggeman_CT <- Bruggeman_CT %>% 
  dplyr::select(Gene_Name)
```

## CTexploreR data for selection pipeline

To characterise the differences between our database and other, we need the 
category we created in CTexploreR. For this, we have the object `all_genes` in 
CTdata that contains the CT analysis for all genes. More info in

Hereunder is what we used for our selection pipeline (coming from 
`make_all_genes_prelim.R` and `130_make_all_genes_and_CT_genes.R` in `CTdata`).

```{r all_genes, message = FALSE, warning = FALSE}
all_genes
```

From there, we filtered based on the testis_specificity ("testis_specific", which
is based on expression in health tissue and scRNA seq info from HPA), 
CCLE_category ("activated") and TCGA_category ("activated" or 
"multimapping_issue") to have our CT genes. Then, when wanting to validate TSS 
manually, we realised that for some genes, reads were not properly aligned to
exons which might reflect a poorly defined transcription in these regions and 
are hence likely unreliable.

Some genes were also characterized as Cancer-Testis preferential genes when 
testis specificity was less stringent

## CTexploreR VS CTdatabase

```{r CTdatabase_ours}
CTdatabase_ours <- Venn(list(CTdatabase = CTdatabase_cleaned$external_gene_name,
                             CTexploreR = CT_genes$external_gene_name))
gp <- VennThemes(compute.Venn(CTdatabase_ours))
gp[["Face"]][["11"]]$fill <-  "mistyrose"
gp[["Face"]][["01"]]$fill <-  "darkseagreen1"
gp[["Face"]][["10"]]$fill <-  "lightsteelblue1"
gp[["Set"]][["Set1"]]$col <-  "paleturquoise4"
gp[["Set"]][["Set2"]]$col <-  "darkseagreen4"
gp[["SetText"]][["Set1"]]$col <-  "paleturquoise4"
gp[["SetText"]][["Set2"]]$col <-  "darkseagreen4"
plot(CTdatabase_ours, gp = gp)

```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_CTdatabase.pdf"))
plot(CTdatabase_ours, gp = gp)
dev.off()
```

We find `r length(CTdatabase_ours@IntersectionSets[["11"]])/nrow(CTdatabase_cleaned)*100`
% of CTdatabase in CTexploreR, which is 
`r length(CTdatabase_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`% of our 
database.

Lost genes analysis

```{r CTdatabase_lost}
CTdatabase_lost <- all_genes %>%
  filter(external_gene_name %in% CTdatabase_ours@IntersectionSets[["10"]])

# 9 Genes are lost because not in any database

table(CTdatabase_lost$testis_specificity)

table(CTdatabase_lost$CT_gene_type)

table(CTdatabase_lost$not_detected_in_somatic_HPA)

table(CTdatabase_lost$TCGA_category)

table(CTdatabase_lost$CCLE_category)

table(CTdatabase_lost$TCGA_category, CTdatabase_lost$CCLE_category)

CTdatabase_lost_upset <- 
  list(`Not testis specific` = 
         filter(CTdatabase_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(CTdatabase_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(CTdatabase_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_CTdatabase <- fromList(CTdatabase_lost_upset)

upset(upset_CTdatabase,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_CTdatabase_lost.pdf"))
upset(upset_CTdatabase,
      text.scale = upset_text_size)
dev.off()
```

`r (nrow(CTdatabase_lost) - table(CTdatabase_lost$testis_specificity)[["testis_specific"]])/nrow(CTdatabase_lost)*100`
% of these genes are not testis specific.

However `r table(CTdatabase_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(CTdatabase_lost) - table(CTdatabase_lost$TCGA_category, CTdatabase_lost$CCLE_category)["activated", "activated"])/nrow(CTdatabase_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

In their analysis, they had characterised gene specificity, some being not 
available, not found in testis, testis-restricted, testis-selective and 
testis/brain-restricted. Let's see how the lost genes qualify as they didn't
mention those were strictly testis specific.

```{r CTdatabase_classification}
CTdatabase_cleaned %>% 
  filter(external_gene_name %in% CTdatabase_lost$external_gene_name) %>% 
  pull(Classification) %>% 
  table()
```

We can see that most of them had no info or were testis-selective (I couldn't
find on website or paper how they selected categories).

## CTexploreR VS omics databases

```{r upset_plot}
core_ours <- Venn(list(Jamin = Jamin_core_CT$Gene, 
                       CTexploreR = CT_genes$external_gene_name))

Wang_ours <- Venn(list(Wang = Wang_CT$external_gene_name, 
                       CTexploreR = CT_genes$external_gene_name))

Carter_ours <- Venn(list(Carter_CT = Carter_CT$Gene_Name, 
                         CTexploreR = CT_genes$external_gene_name))

Bruggeman_ours <- Venn(list(Bruggeman = Bruggeman_CT$Gene_Name,
                            CTexploreR = CT_genes$external_gene_name))

gene_list <- list(CTexploreR = CT_genes$external_gene_name,
                  Carter = Carter_CT$Gene_Name,
                  Jamin = Jamin_core_CT$Gene, 
                  CTatlas = Wang_CT$external_gene_name,
                  Bruggeman = Bruggeman_CT$Gene_Name)

upset_omics <- fromList(gene_list)
upset(upset_omics)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_omics.pdf"))
upset(upset_omics)
dev.off()
```

4 in all, 60 in at least 3 databases

Lost genes analysis

```{r Jamin_lost}
plot(core_ours, gp = gp)

Jamin_lost <- all_genes %>%
  filter(external_gene_name %in% core_ours@IntersectionSets[["10"]])

table(Jamin_lost$testis_specificity)

table(Jamin_lost$CT_gene_type)

table(Jamin_lost$not_detected_in_somatic_HPA)

table(Jamin_lost$TCGA_category)

table(Jamin_lost$CCLE_category)

table(Jamin_lost$TCGA_category, Jamin_lost$CCLE_category)

Jamin_lost_upset <- 
  list(`Not testis specific` = 
         filter(Jamin_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Jamin_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Jamin_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Jamin <- fromList(Jamin_lost_upset)
upset(upset_Jamin,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Jamin.pdf"))
plot(core_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Jamin_lost.pdf"))
upset(upset_Jamin,
      text.scale = upset_text_size)
dev.off()
```

We find `r length(core_ours@IntersectionSets[["11"]])/nrow(Jamin_core_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(core_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Jamin_lost) - table(Jamin_lost$testis_specificity)[["testis_specific"]])/nrow(Jamin_lost)*100`% of 
these genes are not testis specific.

However `r table(Jamin_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Jamin_lost) - table(Jamin_lost$TCGA_category, Jamin_lost$CCLE_category)["activated", "activated"])/nrow(Jamin_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Wang_lost}
plot(Wang_ours, gp = gp)

Wang_lost <- all_genes %>%
  filter(external_gene_name %in% Wang_ours@IntersectionSets[["10"]])

table(Wang_lost$testis_specificity)

table(Wang_lost$CT_gene_type)

table(Wang_lost$not_detected_in_somatic_HPA)

table(Wang_lost$TCGA_category)

table(Wang_lost$CCLE_category)

table(Wang_lost$TCGA_category, Wang_lost$CCLE_category)

Wang_lost_upset <- 
  list(`Not testis specific` = 
         filter(Wang_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Wang_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Wang_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Wang <- fromList(Wang_lost_upset)
upset(upset_Wang,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Wang.pdf"))
plot(Wang_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Wang_lost.pdf"))
upset(upset_Wang,
      text.scale = upset_text_size)
dev.off()
```

We find `r length(Wang_ours@IntersectionSets[["11"]])/nrow(Wang_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Wang_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Wang_lost) - table(Wang_lost$testis_specificity)[["testis_specific"]])/nrow(Wang_lost)*100`% of 
these genes are not testis specific.

However `r table(Wang_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Wang_lost) - table(Wang_lost$TCGA_category, Wang_lost$CCLE_category)["activated", "activated"])/nrow(Wang_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Carter_lost}
plot(Carter_ours, gp = gp)

Carter_lost <- all_genes %>%
  filter(external_gene_name %in% Carter_ours@IntersectionSets[["10"]])

table(Carter_lost$testis_specificity)

table(Carter_lost$CT_gene_type)

table(Carter_lost$not_detected_in_somatic_HPA)

table(Carter_lost$TCGA_category)

table(Carter_lost$CCLE_category)

table(Carter_lost$TCGA_category, Carter_lost$CCLE_category)

Carter_lost_upset <- 
  list(`Not testis specific` = 
         filter(Carter_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Carter_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Carter_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Carter <- fromList(Carter_lost_upset)
upset(upset_Carter,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Carter.pdf"))
plot(Carter_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Carter_lost.pdf"))
upset(upset_Carter,
      text.scale = upset_text_size)
dev.off()
```

We find `r length(Carter_ours@IntersectionSets[["11"]])/nrow(Carter_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Carter_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Carter_lost) - table(Carter_lost$testis_specificity)[["testis_specific"]])/nrow(Carter_lost)*100`% of 
these genes are not testis specific.

However `r table(Carter_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Carter_lost) - table(Carter_lost$TCGA_category, Carter_lost$CCLE_category)["activated", "activated"])/nrow(Carter_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

```{r Bruggeman_lost}
plot(Bruggeman_ours, gp = gp)

Bruggeman_lost <- all_genes %>%
  filter(external_gene_name %in% Bruggeman_ours@IntersectionSets[["10"]])

table(Bruggeman_lost$testis_specificity)

table(Bruggeman_lost$CT_gene_type)

table(Bruggeman_lost$not_detected_in_somatic_HPA)

table(Bruggeman_lost$TCGA_category)

table(Bruggeman_lost$CCLE_category)

table(Bruggeman_lost$TCGA_category, Bruggeman_lost$CCLE_category)

Bruggeman_lost_upset <- 
  list(`Not testis specific` = 
         filter(Bruggeman_lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(Bruggeman_lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(Bruggeman_lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_Bruggeman <- fromList(Bruggeman_lost_upset)
upset(upset_Bruggeman,
      text.scale = upset_text_size)
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/Venn_Bruggeman.pdf"))
plot(Bruggeman_ours, gp = gp)
dev.off()

pdf(file = paste0("figs/upset_Bruggeman_lost.pdf"))
upset(upset_Bruggeman,
      text.scale = upset_text_size)
dev.off()
```

We find `r length(Bruggeman_ours@IntersectionSets[["11"]])/nrow(Bruggeman_CT)*100`
% of CTdatabase in CTexploreR, which is `r length(Bruggeman_ours@IntersectionSets[["11"]])/nrow(CT_genes)*100`
% of our database.

`r (nrow(Bruggeman_lost) - table(Bruggeman_lost$testis_specificity)[["testis_specific"]])/nrow(Bruggeman_lost)*100`% of 
these genes are not testis specific.

However `r table(Bruggeman_lost$CT_gene_type)[["CTP_gene"]]` of these lost 
genes are flagged as Cancer-Testis preferential in our analysis.

`r (nrow(Bruggeman_lost) - table(Bruggeman_lost$TCGA_category, Bruggeman_lost$CCLE_category)["activated", "activated"])/nrow(Bruggeman_lost)*100`
% are not properly activated in tumors and/or cancer cell lines.

## Characterisation of differences with all databases

```{r all_databases}
common <- unique(c(core_ours@IntersectionSets[["11"]], 
                   CTdatabase_ours@IntersectionSets[["11"]], 
                   Wang_ours@IntersectionSets[["11"]], 
                   Carter_ours@IntersectionSets[["11"]],
                   Bruggeman_ours@IntersectionSets[["11"]]))

length(common)

length(common)/dim(CT_genes)[1] * 100

lost_list <- unique(c(core_ours@IntersectionSets[["10"]],
                      CTdatabase_ours@IntersectionSets[["10"]],
                      Wang_ours@IntersectionSets[["10"]],
                      Carter_ours@IntersectionSets[["10"]],
                      Bruggeman_ours@IntersectionSets[["10"]]))

lost <- all_genes %>%
  filter(external_gene_name %in% lost_list)

all_lost_upset <- 
  list(`Not testis specific` = 
         filter(lost,
                testis_specificity != "testis_specific")$external_gene_name,
       `Not tumour activated` = 
         filter(lost,
                (TCGA_category != "activated" &
                  TCGA_category != "multimapping_issue")|
                   CCLE_category != "activated")$external_gene_name,
       `CT preferential` =
         filter(lost,
                CT_gene_type == "CTP_gene")$external_gene_name)

upset_all <- fromList(all_lost_upset)
upset(upset_all,
      text.scale = upset_text_size)

not_specific <- filter(lost, testis_specificity == "not_testis_specific")

GTEX_expression(not_specific$external_gene_name, units = "log_TPM")

somatic_testis <- filter(lost, not_detected_in_somatic_HPA == FALSE)

testis_expression(somatic_testis$external_gene_name, cells = "all")
HPA_cell_type_expression(somatic_testis$external_gene_name)

not_TCGA_activated <- filter(lost, TCGA_category != "activated" & 
                               TCGA_category != "multimapping_issue")

TCGA_expression(not_TCGA_activated$external_gene_name,
                tumor = "all",
                units = "log_TPM")

not_CCLE_activated <- filter(lost, CCLE_category  != "activated")

CCLE_expression(not_CCLE_activated$external_gene_name,
                  type = c("lung", "skin", "colorectal",
                           "gastric", "breast", "head_and_neck"),
                units = "log_TPM")

transcript_prob <- lost %>% 
  filter(testis_specificity == "testis_specific" |
           testis_specificity == "testis_preferential") %>%
  filter(TCGA_category == "activated" | TCGA_category == "multimapping_issue") %>% 
  filter(CCLE_category == "activated") %>% 
  dim()
```

```{r, eval = FALSE, echo = FALSE}
pdf(file = paste0("figs/upset_all_lost.pdf"))
upset(upset_all,
      text.scale = upset_text_size)
dev.off()
```

`r length(common)` genes in our CTexploreR database are found in at least one of
the other database, which represents `r length(common)/dim(CT_genes)[1] * 100`%.

We have lost `r length(lost_list)` genes in total. Among them, 
`r dim(not_specific)[1]/length(lost_list)*100`% are not considered testis 
specific, `r dim(somatic_testis)[1]/length(lost_list)*100`% are expressed in 
somatic cells, `r dim(not_TCGA_activated)[1]/length(lost_list)*100`% 
are not activated in TCGA samples, 
`r dim(not_CCLE_activated)[1]/length(lost_list)*100`% are not activated in CCLE 
cell lines and `r transcript_prob[1]/length(lost_list)*100`% is lost due to 
transcripts problems.

What about new genes in CTexploreR

```{r new, warning = FALSE}
new <- CT_genes %>% 
  filter(!external_gene_name%in%common)

new

table(new$testis_specificity)
table(new$X_linked)
table(new$regulated_by_methylation)
table(new$X_linked, new$regulated_by_methylation)
TCGA_expression(tumor = "all", genes = new$external_gene_name, 
                units = "log_TPM")
TCGA_expression(tumor = "all", 
                genes = filter(new, X_linked & regulated_by_methylation)$external_gene_name, 
                units = "log_TPM")
```

There are `r dim(new)[1]` new CT genes in CTexploreR. These are all testis
specific and mainly on autosomes. Regulation by methylation is the majority of 
them.  There 
is only 11 new "major" CT that are on the X chromosome and regulated by 
methylation. CT45 are not that new.

Expression in tumours doesn't strike that much.

# SessionInfo

```{r}
sessionInfo()
```