### 2024-10-18 Enrichment_unduplicated_family.Rmd
Added enrichment analysis removing duplicates in family

---
title: "Is there still an enrichment of CT genes on the X chromosome if we take duplicated genes in family into account ? "
author: "Julie Devis"
date: "2024-10-18"
output: 
  html_document:
    theme: cosmo
    df_print : paged
    highlight : default
    toc: yes
    number_sections : TRUE
    toc_depth : 4
---

```{r, message = FALSE}
.libPaths("/storage/research/dduv/cbio-lg/rlib/Bioc320", include.site = FALSE)
library(CTexploreR)
library(tidyverse)
```

We've shown that there is an enrichment of CT genes on the X chromosome, and that 
this enrichment was only seen for genes that are MethDep. However, we wonder if
this would still be seen if we take into account the fact that many CT genes are
duplication of each other as they're from the same families.

```{r}
CT_genes <- CT_genes %>% 
  filter(CT_gene_type == "CT_gene")
table(CT_genes$family)
sum(table(CT_genes$family))
table(CT_genes$family, CT_genes$chr)
table(CT_genes$family, CT_genes$regulated_by_methylation)

```

There are 10 families containing 58 genes.

All members of a family are on the same chromosome (only X and Y) and the same 
regulated by methylation status. I can thus only keep one member of each family
and it will be representative of it all. It'll thus remove 46 genes from my 
list (58 - 10 which is 1 of each family kept)

```{r}
CT_no_duplicated_fam <- CT_genes %>% 
  filter(!duplicated(family) | is.na(family)) %>% 
  mutate(Regulation = case_when(regulated_by_methylation ~ "Methylation",
                                !regulated_by_methylation ~ "Not methylation")) %>%
  mutate(Chr = case_when(X_linked ~ "X-linked",
                         !X_linked ~ "not X")) 
dim(CT_no_duplicated_fam)

CT_no_duplicated_fam$chr <- factor(CT_no_duplicated_fam$chr, levels = c(1:22, "X", "Y"))

fig_X_enrichment <- ggplot(CT_no_duplicated_fam %>% 
                             filter(!is.na(regulated_by_methylation)), 
                           aes(x = chr, fill = Regulation)) +
  geom_bar(stat = 'count') +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = c("Methylation" = "indianred2",
                               "Not methylation"= "cyan4"))+
  theme_bw() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.2, 0.85),
        #aspect.ratio = 1.5,
        panel.grid = element_blank())

fig_X_enrichment

table(CT_no_duplicated_fam$chr)

CT_no_duplicated_fam_rep <- CT_no_duplicated_fam %>% 
  filter(!is.na(regulated_by_methylation)) %>% 
  group_by(Regulation, Chr) %>%
  dplyr::count() %>% 
  pivot_wider(names_from = Regulation, values_from = n)

CT_no_duplicated_fam_rep

chisq.test(CT_no_duplicated_fam_rep[,-1])

```

We can still see that there is more genes on the X chromosome than on others.
However, in comparison with the analysis not removing duplicated families, 
the MethDep repartition between X and not X is different, as there is now more 
genes that are not X (37 out of the initial 39) than those that are X-linked 
(27 out of the 69). However, pvalue is still significative.

Could also be done by autosomes/X

```{r}
CT_genes_X_linked <- CT_no_duplicated_fam %>% 
  mutate(chr = case_when(chr == "X" ~ "X-linked",
                           chr == "Y" ~ "Y-linked",
                           !chr %in% c("X", "Y") ~ "Autosome")) %>% 
  filter(!is.na(regulated_by_methylation)) %>% 
  group_by(Regulation, chr) %>%
  dplyr::count() %>% 
  pivot_wider(names_from = Regulation, values_from = n)
CT_genes_X_linked
```

What does it give if we normalise according to the number of gene per chromosome ?

```{r}
n_genes_per_chr <- all_genes %>% 
  filter(chr != "MT") %>% 
  group_by(chr) %>% 
  summarize(n = n()) %>% 
  mutate(chr = factor(chr, levels = c(1:22, "X", "Y"))) %>% 
  arrange(chr)

n_genes <- n_genes_per_chr %>% 
  filter(!chr %in% "MT") %>% 
  pull(n) %>% 
  sum()
n_genes_autosome <- n_genes_per_chr %>% 
  filter(chr %in% 1:22) %>% 
  pull(n) %>% 
  sum()
n_genes_X <- n_genes_per_chr %>% 
  filter(chr %in% "X") %>% 
  pull(n) 
n_genes_Y <- n_genes_per_chr %>% 
  filter(chr %in% "Y") %>% 
  pull(n)

n_genes_per_chr_autosome_X_Y <- tibble("chr" = c("Autosome", "X-linked", "Y-linked"), 
       n = c(n_genes_autosome, n_genes_X, n_genes_Y))

ggplot(n_genes_per_chr, aes(x = chr, y = n)) +
  geom_col() +
  xlab("Chromosome") +
  ylab("Gene number") +
  scale_fill_manual(values = c("Methylation" = "indianred2",
                               "Not methylation"= "cyan4")) +
  theme_bw() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.2, 0.85),
        #aspect.ratio = 1.5,
        panel.grid = element_blank()) +
  ggtitle("Total number of genes per chromosome")

n_genes_per_chr_autosome_X_Y <- n_genes_per_chr_autosome_X_Y %>% 
  mutate(proba = n / n_genes)
n_genes_per_chr_autosome_X_Y

CT_genes_Met <- colSums(CT_genes_X_linked[, -1])["Methylation"]
CT_genes_not_Met <- colSums(CT_genes_X_linked[, -1])["Not methylation"]
CT_genes_X_Y_observed_expected <- CT_genes_X_linked %>% 
  left_join(n_genes_per_chr_autosome_X_Y %>% 
              dplyr::select(-n)) %>% 
  mutate(Expected_MET_if_random = (CT_genes_Met * proba),
         Expected_NOT_MET_if_random = (CT_genes_not_Met * proba)) %>% 
  dplyr::select(chr, Methylation, Expected_MET_if_random, `Not methylation`, Expected_NOT_MET_if_random)
CT_genes_X_Y_observed_expected

chisq.test(CT_genes_X_Y_observed_expected[
  CT_genes_X_Y_observed_expected$chr %in% c("Autosome", "X-linked"), c("Methylation", "Expected_MET_if_random")]) 
```

CT genes regulated by met are enriched on X when taking into account the number 
of genes per chromosome, verified statistically.