### 2025-03-25 3.CT_in_tumors.Rmd

---
title: "CT genes in tumors"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: false
---

```{r, echo = FALSE, Results = 'asis'}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library("tidyverse")
library("CTexploreR")
library("ComplexHeatmap")
library("circlize")
library("SummarizedExperiment")
```

#

```{r}
load("~/cluster/Packages/CTdata/eh_data/all_genes.rda")

CT_genes_all <- all_genes

CT_genes_types_all <- all_genes %>% 
  dplyr::rename(gene = external_gene_name) %>% 
  mutate(Regulation = case_when(regulated_by_methylation ~ "Methylation",
                                !regulated_by_methylation ~ "Not methylation")) %>%
  mutate(Chr = case_when(X_linked ~ "X-linked",
                         !X_linked ~ "not X")) %>% 
  mutate(chr_met = case_when(regulated_by_methylation & X_linked ~ "X_met",
                             !regulated_by_methylation & X_linked ~ "X_not_met",
                             regulated_by_methylation & !X_linked ~ "not_X_met",
                             !regulated_by_methylation & !X_linked ~ "not_X_not_met")) %>% 
  mutate(GTEX_specificity = case_when(GTEX_category == "testis_specific" ~ "Testis specific",
                                        GTEX_category == "testis_preferential" ~ "Testis preferential",
                                        GTEX_category == "lowly_expressed" ~ "Undetectable in GTEX")) %>% 
  mutate(GTEX_specificity = factor(GTEX_specificity, 
                                     levels = c("Testis specific",  "Undetectable in GTEX", "Testis preferential"))) %>% 
  mutate(testis_specificity = case_when(testis_specificity == "testis_specific" ~ "Testis specific",
                                        testis_specificity == "testis_preferential" ~ "Testis preferential")) %>%
  mutate(testis_specificity = factor(testis_specificity, levels = c("Testis specific", "Testis preferential"))) %>% 
  mutate(chr_met = factor(chr_met, 
                          levels = c("X_met", "not_X_met", "X_not_met", "not_X_not_met"))) %>% 
  mutate(met_3_groups = case_when(regulated_by_methylation & X_linked ~ "Methylation X-linked",
                                   regulated_by_methylation & !X_linked ~ "Methylation not X-linked",
                             !regulated_by_methylation ~ "Not Methylation")) %>% 
  mutate(met_3_groups = factor(met_3_groups, 
                          levels = c("Methylation X-linked", "Methylation not X-linked", "Not Methylation")))

CT_genes <- CT_genes_all %>%
  filter(CT_gene_type == "CT_gene")
CT_genes_types <- CT_genes_types_all %>% 
  filter(CT_gene_type == "CT_gene")

CTP_genes <- CT_genes_all %>%
  filter(CT_gene_type == "CTP_gene")
CTP_genes_types <- CT_genes_types_all %>% 
  filter(CT_gene_type == "CTP_gene")
```

```{r}
legends_param <- list(
  labels_gp = gpar(col = "black", fontsize = 4),
  title_gp = gpar(col = "black", fontsize = 5, fontface = "bold"),
  simple_anno_size = unit(0.2, "cm"),
  row_names_gp = gpar(fontsize = 3),
  annotation_name_side = "left",
  border = FALSE,
  border_gp = gpar(lwd = 0.2),
  grid_width = unit(0.2, "cm"),
  grid_height = unit(0.05, "cm"),
  legend_height = unit(1, "cm"),
  use_raster = TRUE,
               raster_device = "CairoPNG",
               raster_quality = 10)

legend_colors <- c("#5E4FA2", "#3288BD", "#66C2A5", "#ABDDA4", "#E6F598",
"#FFFFBF", "#FEE08B", "#FDAE61", "#F46D43", "#D53E4F",
"#9E0142")

TCGA_colors <- c(
"BRCA" = "midnightblue", "COAD" = "darkorchid2",
"ESCA" = "gold", "HNSC" = "deeppink2",
"LUAD" = "seagreen", "LUSC" = "seagreen3",
"SKCM" = "red3")

CCLE_colors <- c(
"Lung" = "seagreen3", "Skin" = "red3",
"Bile_Duct" = "mediumpurple1", "Bladder" = "mistyrose2",
"Colorectal" = "plum", "Lymphoma" = "steelblue1",
"Uterine" = "darkorange4", "Myeloma" = "turquoise3",
"Kidney" = "thistle4",
"Pancreatic" = "darkmagenta", "Brain" = "palegreen2",
"Gastric" = "wheat3", "Breast" = "midnightblue",
"Bone" = "sienna1", "Head_and_Neck" = "deeppink2",
"Ovarian" = "tan3", "Sarcoma" = "lightcoral",
"Leukemia" = "steelblue4", "Esophageal" = "khaki",
"Neuroblastoma" = "olivedrab1")

DAC_colors <- c(
"B2-1" = "olivedrab2", "HCT116" = "lightcoral",
"HEK293T" = "seagreen3", "HMLER" = "mediumpurple1",
"IMR5-75" = "deeppink2", "NCH1681" = "steelblue2",
"NCH612" = "red3", "TS603" = "darkmagenta")

DAC_treatment_colors <- c("CTL" = "royalblue4", "5-aza" = "maroon3")

Regulation_colors <- c("Methylation" = "indianred2",
"Not methylation"= "cyan4")

MethDep_colors <- c("MethDep" = "indianred2",
"Non-MethDep"= "cyan4")

chr_met_colors = c("X_met" = "mediumorchid2",
"X_not_met" = "mediumseagreen",
"not_X_met"= "darkorange1",
"not_X_not_met" = "dodgerblue1")

Chr_colors = c("X-linked" = "deeppink3", "not X" = "lightgreen")
Chr_colors_with_Y = c("X-linked" = "deeppink3", "Autosome" = "lightgreen", "Y-linked" = "black")

GTEX_specificity_colors = c("Testis specific" = "green",
                              "Testis preferential" = "purple",
                              "Undetectable in GTEX" = "gray")

Testis_specificity_colors = c("Testis specific" = "mediumpurple2",
"Testis preferential" = "goldenrod1",
"Undetectable in GTEX" = "gray")

# Testis_specificity_colors = c("Testis specific" = "green",
#                               "Testis preferential" = "purple")

testis_colors <- c(
"SSC" = "floralwhite", "Spermatogonia" = "moccasin",
"Early_spermatocyte" = "gold",
"Late_spermatocyte" = "orange",
"Round_spermatid" = "red2",
"Elongated_spermatid" = "darkred", "Sperm" = "purple",
"Sertoli" = "gray",
"Leydig" = "cadetblue2", "Myoid" = "springgreen3",
"Macrophage" = "gray10",
"Endothelial" = "steelblue")

Fetal_cell_colors = c( "F_PGC" = "pink",
"F_GC" = "pink3",
"F_oogonia" = "palevioletred3",
"F_oocyte" = "mediumorchid4",
"M_PGC" = "lightblue1",
"M_GC" = "steelblue3",
"M_pre_spermatogonia" = "royalblue3")

oocytes_colors <- c("Growing oocytes" = "mediumpurple1",
# "Fully grown oocytes" ="lightgreen",
# "Metaphase I" = "mediumvioletred",
"Metaphase II" = "mediumpurple4")

sex_colors <- c("F" = "palevioletred1", "M"= "skyblue1")

stage_colors <- c( "pre-meiotic" = "peachpuff",
"meiotic"= "lightsalmon",
"post-meiotic" = "maroon4",
"somatic" = "gray")

CT_genes <- CT_genes %>% filter(CT_gene_type == "CT_gene")
```

# Activation frequencies / mean expression

```{r}
####################################################################
### TCGA
####################################################################
load("~/cluster/Packages/CTdata/eh_data/TCGA_TPM.rda")
TCGA <- TCGA_TPM
TCGA <- TCGA[, TCGA$shortLetterCode != "NT"]
TCGA_Subtype <- c("TCGA-SKCM", "TCGA-LUAD", "TCGA-HNSC")
TCGA <- TCGA[, TCGA$project_id %in% TCGA_Subtype]

tumors <- unique(TCGA$project_id)

thr_pos <- 1

prop_of_pos_tum <- as_tibble(assay(TCGA), rownames = "ensembl_gene_id") %>% 
  filter(ensembl_gene_id %in% CT_genes$ensembl_gene_id) %>% 
  pivot_longer(names_to = "barcode", values_to = "TPM", -ensembl_gene_id) %>% 
  mutate(activated = ifelse(TPM >= thr_pos, "n_pos", "n_neg")) %>% 
  left_join(as_tibble(colData(TCGA), rownames = "barcode") %>% 
              dplyr::select(barcode, project_id)) %>% 
  group_by(ensembl_gene_id, activated, project_id) %>% 
  dplyr::summarize(n = n()) %>% 
  pivot_wider(names_from = activated, values_from = n) %>% 
  mutate(proportion_of_positive_tum = n_pos * 100 / (n_pos + n_neg))

mean_in_pos <- as_tibble(assay(TCGA), rownames = "ensembl_gene_id") %>% 
  filter(ensembl_gene_id %in% CT_genes$ensembl_gene_id) %>%  
  pivot_longer(names_to = "barcode", values_to = "TPM", -ensembl_gene_id)   %>% 
  mutate(activated = ifelse(TPM >= thr_pos, "activated", "not_activated")) %>% 
  filter(activated == "activated") %>% 
  left_join(as_tibble(colData(TCGA), rownames = "barcode") %>% 
              dplyr::select(barcode, project_id)) %>% 
  group_by(ensembl_gene_id, project_id) %>% 
  summarise(mean_log = log1p(mean(TPM)),
            mean_TPM = mean(TPM))

mean_and_prop_TCGA <- prop_of_pos_tum %>% 
  left_join(mean_in_pos) %>% 
  right_join(CT_genes) %>% 
  dplyr::select(ensembl_gene_id, project_id, mean_log, proportion_of_positive_tum, regulated_by_methylation, X_linked) %>% 
  mutate(source = "TCGA tumor samples")

####################################################################
### CCLE data
####################################################################
load("~/cluster/Packages/CTdata/eh_data/CCLE_data.rda")
CCLE <- CCLE_data
squamous_cell_carcinomas <- unique(grep("Squamous Cell Carcinoma", 
                                        x = CCLE$Subtype, value = TRUE))

CCLE_Subtype <- c("Melanoma", 
                 "Non-Small Cell Lung Cancer (NSCLC), Adenocarcinoma", 
                 squamous_cell_carcinomas)

CCLE <- CCLE[, !is.na(CCLE$Subtype) & CCLE$Subtype %in% CCLE_Subtype]
CCLE$type <- NA
CCLE$type[CCLE$Subtype == "Melanoma" & CCLE$primary_disease == "Skin Cancer"] <- "CCLE-SKCM"
CCLE$type[CCLE$Subtype == "Non-Small Cell Lung Cancer (NSCLC), Adenocarcinoma" & 
            CCLE$primary_disease == "Lung Cancer"] <- "CCLE-LUAD"
CCLE$type[CCLE$Subtype == "Non-Small Cell Lung Cancer (NSCLC), Squamous Cell Carcinoma" & 
            CCLE$primary_disease == "Lung Cancer"] <- "CCLE-LUSC"
CCLE$type[CCLE$Subtype %in% squamous_cell_carcinomas & 
            CCLE$primary_disease == "Head and Neck Cancer"] <- "CCLE-HNSC"

CCLE <- CCLE[, !is.na(CCLE$type) & CCLE$type != "CCLE-LUSC"]
CCLE$type <- factor(CCLE$type, levels = c("CCLE-SKCM", 
                                          "CCLE-LUAD", 
                                          #"CCLE-LUSC", 
                                          "CCLE-HNSC"))

prop_of_pos_tum <- as_tibble(assay(CCLE), rownames = "ensembl_gene_id") %>% 
    filter(ensembl_gene_id %in% CT_genes$ensembl_gene_id) %>%  
    pivot_longer(names_to = "cell", values_to = "TPM", -ensembl_gene_id) %>% 
    mutate(activated = ifelse(TPM >= thr_pos, "n_pos", "n_neg")) %>% 
    left_join(as_tibble(colData(CCLE), rownames = "cell") %>% 
                dplyr::select(cell, Subtype, lineage_subtype, type)) %>% 
    group_by(ensembl_gene_id, activated, type) %>% 
    dplyr::summarize(n = n()) %>% 
    pivot_wider(names_from = activated, values_from = n, values_fill = 0) %>% 
    mutate(proportion_of_positive_tum = n_pos * 100 / (n_pos + n_neg))

mean_in_pos <- as_tibble(assay(CCLE), rownames = "ensembl_gene_id") %>% 
  filter(ensembl_gene_id %in% CT_genes$ensembl_gene_id) %>%   
  pivot_longer(names_to = "cell", values_to = "TPM", -ensembl_gene_id)  %>% 
  mutate(activated = ifelse(TPM >= thr_pos, "activated", "not_activated")) %>% 
  filter(activated == "activated") %>% 
  left_join(as_tibble(colData(CCLE), rownames = "cell") %>% 
              dplyr::select(cell, Subtype, type)) %>% 
  group_by(ensembl_gene_id, type) %>% 
  summarise(mean_log = log1p(mean(TPM)),
            mean_TPM = mean(TPM))

mean_and_prop_CCLE <- prop_of_pos_tum %>% 
  left_join(mean_in_pos) %>% 
  right_join(CT_genes) %>% 
  dplyr::select(ensembl_gene_id, type, mean_log, proportion_of_positive_tum, regulated_by_methylation, X_linked) %>% 
  mutate(source = "CCLE tumor cell lines")

####################################################################
### combine TCGA and CCLE data
####################################################################

bilan <- rbind(mean_and_prop_CCLE, 
               mean_and_prop_TCGA %>% 
                 dplyr::rename(type = project_id))

bilan <- bilan %>% 
  mutate(type = factor(type, levels = c("CCLE-SKCM", 
                                        "CCLE-LUAD", 
                                        #"CCLE-LUSC", 
                                        "CCLE-HNSC",
                                        "TCGA-SKCM", 
                                        "TCGA-LUAD", 
                                        #"TCGA-LUSC", 
                                        "TCGA-HNSC"))) %>% 
  dplyr::rename("Positive tumors (%)" = proportion_of_positive_tum,
                'Mean expression\nin positive tumors' = mean_log)

library(ggh4x)

strip <- strip_themed(background_x = elem_list_rect(fill = c(rep("gray95", 3), rep("gray80", 3))))

fig <- bilan %>% 
  mutate(type = gsub("-", x = type, replace = "\n")) %>% 
  mutate(type = factor(type, 
                       levels =  c("CCLE\nSKCM", 
                                   "CCLE\nLUAD", 
                                    #"CCLE\nLUSC", 
                                   "CCLE\nHNSC",
                                   "TCGA\nSKCM", 
                                   "TCGA\nLUAD", 
                                    #"TCGA\nLUSC", 
                                   "TCGA\nHNSC" ))) %>% 
  mutate("CG gene" = case_when(regulated_by_methylation  ~ "MethDep",
                               !regulated_by_methylation ~ "Non-MethDep")) %>% 
  pivot_longer(names_to = "metric", values_to = "value", 
               -c(ensembl_gene_id,  `CG gene`, type, regulated_by_methylation, X_linked, source)) %>% 
  filter(!is.na(`CG gene`)) %>% 
  ggplot(aes(x = `CG gene`, y = value, color = `CG gene`)) +
  geom_jitter(size = 0.2, alpha = .8) +
  scale_colour_manual(values = MethDep_colors) +
  geom_boxplot(alpha = 0, color = "black", linewidth = 0.4) +
  facet_grid2(metric ~ type, scales = "free", switch = "y", strip = strip) +
  xlab('') +
  theme_bw() +
  theme(#aspect.ratio = 0.8,
        legend.position = "right",
        axis.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.5, 'cm'),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        axis.text.x = element_text(size = 0),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_line(size = 0),
        #axis.ticks.y = element_line(size = 0),
        axis.title = element_text(size = 0),
        strip.text.x = element_text(size = 10),
        strip.text.y = element_text(size = 12),
        strip.background.y =  element_blank()) +
  scale_y_continuous(position = "right")

pdf(file = "../figs/Fig5B_CG_in_CCLE_TCGA.pdf", width = 6, height = 4)
fig
dev.off()

fig
```

## t-tests

```{r}
tumors <- levels(bilan$type)
tum <- tumors[1]
prop_of_positive_tum_pval <- t.test(bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Positive tumors (%)") %>% 
                filter(regulated_by_methylation) %>% 
                pull(value),
              bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Positive tumors (%)") %>% 
                filter(regulated_by_methylation) %>% 
                pull(value), alternative = "greater")$p.value

mean_log_pval = t.test(bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Mean expression\nin positive tumors") %>% 
                filter(regulated_by_methylation) %>% 
                pull(value),
              bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Mean expression\nin positive tumors") %>% 
                filter(!regulated_by_methylation) %>% 
                pull(value), alternative = "greater")$p.value

t_tests <- tibble(Tum = tum, prop_of_positive_tum_pval = prop_of_positive_tum_pval, mean_log_pval = mean_log_pval)

for (tum in tumors[-1]){

  tmp1 = t.test(bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Positive tumors (%)") %>% 
                filter(regulated_by_methylation) %>% 
                pull(value),
              bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Positive tumors (%)") %>% 
                filter(!regulated_by_methylation) %>% 
                pull(value), alternative = "greater")$p.value

  tmp2 = t.test(bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Mean expression\nin positive tumors") %>% 
                filter(regulated_by_methylation) %>% 
                pull(value),
              bilan %>% 
                pivot_longer(names_to = "metric", values_to = "value", 
                             -c(ensembl_gene_id,  type, regulated_by_methylation, X_linked, source)) %>% 
                filter(!is.na(regulated_by_methylation)) %>% 
                filter(type == tum, metric == "Mean expression\nin positive tumors") %>% 
                filter(!regulated_by_methylation) %>% 
                pull(value), alternative = "greater")$p.value

  tmp <- tibble(Tum = tum, prop_of_positive_tum_pval = tmp1,  mean_log_pval = tmp2)
  t_tests <- rbind(t_tests, tmp)
}

t_tests$prop_of_positive_tum_padjusted <- p.adjust(t_tests$prop_of_positive_tum_pval, method = "BH")
t_tests$Mean_expression_padjusted <- p.adjust(t_tests$mean_log_pval, method = "BH")
DT::datatable(t_tests %>% dplyr::select(Tum, prop_of_positive_tum_padjusted, Mean_expression_padjusted))
```

=> Always significant (except the `proportion of positive tumors` in TCGA-LUAD)

```{r}
strip <- strip_themed(background_x = elem_list_rect(fill = c("gray95", "gray80")))

bilan_all_types <- bilan %>% 
         filter(!is.na(regulated_by_methylation)) %>% 
         mutate("CG gene" = case_when(regulated_by_methylation ~ "MethDep",
                               !regulated_by_methylation ~ "Non-MethDep")) %>% 
       filter(!is.na(regulated_by_methylation)) %>% 
  group_by(ensembl_gene_id, source) %>% 
  summarize(`Mean expression\nin positive tumors` = mean(`Mean expression\nin positive tumors`, na.rm = TRUE),
            `Positive tumors (%)` = mean(`Positive tumors (%)`, na.rm = TRUE)) %>% 
  left_join(bilan %>% filter(!is.na(regulated_by_methylation)) %>% 
         mutate("CG gene" = case_when(regulated_by_methylation ~ "MethDep",
                               !regulated_by_methylation ~ "Non-MethDep")) %>% 
           dplyr::select(ensembl_gene_id, "CG gene", "X_linked") %>% unique())
```

```{r}

selected_genes_in_TCGA <- c("MAGEA3", "MAGEA6", "MAGEA4", "MAGEA1", "MAGEA12", 
                            "MAGEC2", "CTAG2", "CBY2", "PAGE2", "PAGE5, GAGE2A", 
                            "SPANXB1", "PAGE1", "LINC01518", "BRDT", #"CT45A10", 
                            "SSX1", "MAGEC2", "MAGEB2", "MAGEA10")

selected_genes_in_CCLE <- c("MAGEA3", "MAGEA6", "MAGEA4", #"MAGEA1",
                            "TRIML2", "LINC02864", 
                            "MAGEA12",
                            "MAGEA2", "SPANXB1","LIN1518",
                            "XAGE1A", "XAGE1B", "CTAG2", "CT45A1", "GAGE2A", 
                            "PAGE2", "DSCR8", "MAGEB2", #"VCY", "SSX2B", "TUBA3C",
                            "GAGE12E")

fig <- ggplot(bilan_all_types %>% left_join(CT_genes %>% 
                                              dplyr::select(ensembl_gene_id, external_gene_name) %>% 
                                              dplyr::rename(gene = external_gene_name)) %>% 
                mutate(source = factor(source, 
                       levels =  c("TCGA tumor samples", "CCLE tumor cell lines"))) %>% 
         mutate(Location = case_when(X_linked ~ "X-linked",
                                       !X_linked ~ "Not X")),
         aes(x = `Mean expression\nin positive tumors`, y = `Positive tumors (%)`,
             label = gene, colour = `CG gene`, shape = Location)) +
  geom_point(size = 2) +
  scale_colour_manual(values = MethDep_colors) +
  facet_wrap2(~ source, ncol = 2, strip = strip) +
  ggrepel::geom_text_repel(
    data = bilan_all_types %>% 
      left_join(CT_genes %>% 
                  dplyr::select(ensembl_gene_id, external_gene_name) %>% 
                  dplyr::rename(gene = external_gene_name)) %>% 
      mutate(Location = case_when(X_linked ~ "X-linked",
                                  !X_linked ~ "Not X")) %>%
      filter(source == "TCGA tumor samples" & gene %in% selected_genes_in_TCGA) ,
    aes(x = `Mean expression\nin positive tumors`, y = `Positive tumors (%)`,
        label = gene), size = 2.5, color = "black", min.segment.length = 2,
    max.overlaps = 30) + #, direction = "x") +
  ggrepel::geom_text_repel(
    data = bilan_all_types %>% 
      left_join(CT_genes %>% 
                  dplyr::select(ensembl_gene_id, external_gene_name) %>% 
                  dplyr::rename(gene = external_gene_name)) %>% 
      mutate(Location = case_when(X_linked ~ "X-linked",
                                  !X_linked ~ "Not X")) %>%
      filter(source == "CCLE tumor cell lines" & gene %in% selected_genes_in_CCLE) ,
    aes(x = `Mean expression\nin positive tumors`, y = `Positive tumors (%)`,
        label = gene), size = 2.5, color = "black", min.segment.length = 2,
    max.overlaps = 30) + #, direction = "x") +
  xlab("Mean expression log(TPM + 1) in positive tumors") +
  ylab("Proportion of positive tumors (%)") +

  theme_bw() +
  theme(legend.position = "right",
        #aspect.ratio = 0.8,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        #legend.key.size = unit(0.3, 'cm'),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        axis.ticks = element_line(size = 0.1),
        axis.title = element_text(size = 12),
        strip.text = element_text(size = 12))

pdf(file = "../figs/Fig5C_CG_in_CCLE_TCGA_by_chr.pdf", width = 7, height = 3)
fig
dev.off()

fig

CCLE_data <- CCLE
TCGA_data <- TCGA
```

# CT-met cluster together

```{r}
legends_param <- list(
  labels_gp = gpar(col = "black", fontsize = 6),
  title_gp = gpar(col = "black", fontsize = 6),
  simple_anno_size = unit(0.8, "cm"),
  border_gp = gpar(lwd = 0.01),
  row_names_gp = gpar(fontsize = 1),
  annotation_name_side = "none")

tum_heatmap <- function(genes = NULL,
                               CCLE_Subtype = c("CCLE-SKCM", "CCLE-LUAD", 
                                            #"CCLE-LUSC",  
                                            "CCLE-HNSC"),
                               TCGA_Subtype = c("TCGA-SKCM", "TCGA-LUAD", 
                                            #"TCGA-LUSC",  
                                            "TCGA-HNSC"),
                               h_width = 5,
                               h_height = 10,
                               downsample_to_n = 450,
                               fontsize = 4, 
                               clust_method = "ward.D",
                               split_rows_by = "regulation", 
                               show_annotation_legend = TRUE,
                               clustRow = FALSE,
                               clustCol = TRUE){
  # CCLE data
  ## downsample tumors to have the same number of tumors in each type
  CCLE <- CCLE_data[, CCLE_data$type == CCLE_Subtype[1]]
  CCLE <- CCLE[, 1:min(ncol(CCLE), downsample_to_n)]

  for (tum in CCLE_Subtype[-1]){
    tmp <- CCLE_data[, CCLE_data$type == tum]
    tmp <- tmp[, 1:min(ncol(tmp), downsample_to_n)]
    CCLE <- cbind(CCLE, tmp)
  }

  CCLE$type <- gsub("-", x = CCLE$type, replacement = "\n")
  CCLE$type <- factor(CCLE$type, levels = c("CCLE\nSKCM", "CCLE\nLUAD", #"CCLE\nLUSC", 
                                            "CCLE\nHNSC"))

  colData(CCLE) <- colData(CCLE)[, c(14, 30)]

  # TCGA data
  TCGA <- TCGA_data

  ## downsample tumors to have the same number of tumors in each type
  TCGA <- TCGA_data[, TCGA_data$project_id == TCGA_Subtype[1]]
  TCGA <- TCGA[, 1:min(ncol(TCGA), downsample_to_n)]

  for (tum in TCGA_Subtype[-1]){
    tmp <- TCGA_data[, TCGA_data$project_id == tum]
    tmp <- tmp[, 1:min(ncol(tmp), downsample_to_n)]
    TCGA <- cbind(TCGA, tmp)
  }

  TCGA$type <- TCGA$project_id
  TCGA$type <- gsub("-", x = TCGA$type, replacement = "\n")
  TCGA$type <- factor(TCGA$type, levels = c("TCGA\nSKCM", "TCGA\nLUAD", #"TCGA\nLUSC", 
                                            "TCGA\nHNSC"))

  #names(colData(TCGA))
  colData(TCGA) <- colData(TCGA)[, c(56, 66)]
  #TCGA$project_id <- gsub(pattern = "TCGA-", x = TCGA$project_id, replacement = '')
  my_genes <- genes[genes %in% c(rowData(CCLE)$external_gene_name, rowData(TCGA)$external_gene_name)]
  my_ensembl <- CT_genes %>% filter(external_gene_name %in% my_genes) %>% pull(ensembl_gene_id)

  print(length(my_genes))

  names(colData(TCGA)) <- c("Subtype", "type")
  tum_SE <- cbind(CCLE[my_ensembl,], TCGA[my_ensembl,])

  df_col <- tibble(sample = colnames(tum_SE), type = tum_SE$type)

  df_col <- df_col %>% arrange(type)

  tum_SE <- tum_SE[, df_col$sample]
  split_by <- tum_SE$type

  tum_col <- c(rep("gray95", 4), rep("gray80", 4))
  names(tum_col) <- levels(tum_SE$type)

  ## Use gene names instead of ENSEMBL IDs
  mat <- SummarizedExperiment::assay(tum_SE)
  rownames(mat) <- rowData(tum_SE)$external_gene_name
  mat <- log1p(mat)

  gene_type <- CT_genes %>% 
    filter(external_gene_name %in% genes) %>% 
    mutate(gene = factor(external_gene_name, levels = genes)) %>% 
    mutate(Regulation = ifelse(regulated_by_methylation, "MethDep", "Non\nMethDep")) %>% 
     mutate(Regulation = factor(Regulation, levels = c("MethDep", "Non\nMethDep"))) %>% 
    arrange(gene)

 if (clustRow) gene_type <- gene_type %>% arrange(Regulation)
 rowSplit <- gene_type$Regulation

  heatmap_colors <- colorRamp2(
    seq(0, 5, length = 11),                 
    legend_colors)

  Heatmap(mat[as.vector(gene_type$gene), ],
          top_annotation = HeatmapAnnotation(
            type = anno_block(gp = gpar(fill = c(rep("gray95", 3), rep("gray80", 3)), lwd = 0.2),
                              height = unit(c(0.7), "cm"),
                              labels = levels(df_col$type), 
                              labels_gp = gpar(col = "black", fontsize = 8))),
          left_annotation = rowAnnotation(
            met = anno_block(gp = gpar(fill = MethDep_colors, lwd = 0.2),
                             width = unit(c(0.8), "cm"),
                             labels = c("MethDep", "Non\nMethDep"), 
                             labels_gp = gpar(col = "white", fontsize = 10))),
          name = "logTPM",
          #column_title = paste0("CCLE tumor cell lines"),
          #column_title_gp = gpar(fontsize = 7, face = "bold"),
          clustering_method_rows = clust_method,
          clustering_method_columns = clust_method,
          column_split = split_by,
          row_gap = unit(c(1), "mm"),
          col = heatmap_colors,
          cluster_rows = clustRow,
          cluster_columns = clustCol,
          cluster_column_slices = FALSE,
          cluster_row_slices = FALSE,
          show_row_names = FALSE,
          show_column_names = FALSE,
          show_row_dend = FALSE,
          show_column_dend = FALSE,
          column_names_gp = gpar(fontsize = 6),
          column_names_centered = TRUE,
          row_names_gp = gpar(fontsize = fontsize),
          row_names_side = "left",
          border = TRUE,
          border_gp = gpar(lwd = 0.5),
          column_names_side = c("top"),
          column_names_rot = 90,
          row_title_gp = gpar(fontsize = 0),
          column_title_gp = gpar(fontsize = 0),
          #top_annotation = column_ha_tumor,
          #right_annotation = rAnnot,
          #left_annotation = rAnnot,
          row_split = rowSplit,
          width = unit(h_width, "cm"),
          height = unit(h_height, "cm"),
          use_raster = TRUE,
          raster_device = "CairoPNG",
          raster_quality = 10,
          heatmap_legend_param = legends_param,
          show_heatmap_legend = TRUE)
}
```

```{r}
met_genes <- CT_genes %>% 
  filter(regulated_by_methylation) %>% 
  pull(external_gene_name)

not_met_genes <- CT_genes %>% 
  filter(!regulated_by_methylation) %>% 
  pull(external_gene_name)

h <- tum_heatmap(c(met_genes, not_met_genes), 
                         downsample_to_n = 50, 
                         h_width = 7,
                         h_height = 6,
                         fontsize = 0, 
                   clust_method = "ward.D",
                   clustRow = TRUE,
                   #show_row_names = TRUE,
                   show_annotation_legend = TRUE,
                   split_rows_by = "regulation")
h
pdf(file = "../figs/Fig5A_CG_clustering_in_tumors.pdf")#, width = 7, height = 3)
h
dev.off()

```

# Activation score distribution

- Activated when TPM >= 1

- Activation score :

    number of methDep gene activated / number of methDep gene

    number of Non-methDep gene activated / number of Non-methDep gene

```{r}
### in CCLE
CCLE <- CCLE_data

keep <- rowSums(assay(CCLE)) > 0
CCLE <- CCLE[keep, ]

met_genes <- CT_genes %>% 
  filter(ensembl_gene_id %in% rownames(CCLE)) %>% 
  filter(regulated_by_methylation) %>% 
  pull(ensembl_gene_id)

not_met_genes <- CT_genes %>% 
  filter(ensembl_gene_id %in% rownames(CCLE)) %>% 
  filter(!regulated_by_methylation) %>% 
  pull(ensembl_gene_id)

CCLE <- CCLE[c(met_genes,not_met_genes), ]
binary_mat_methDep <- assay(CCLE) >= 1
methDep_score <- colSums(binary_mat_methDep) / length(met_genes)
CCLE$methDep_score <- methDep_score

binary_mat_NonMethDep <- assay(CCLE[not_met_genes,]) >= 1
NonMethDep_score <- colSums(binary_mat_NonMethDep) / length(not_met_genes)
CCLE$NonMethDep_score <- NonMethDep_score

### in TCGA
TCGA <- TCGA_data

keep <- rowSums(assay(TCGA)) > 0
TCGA <- TCGA[keep, ]

met_genes <- CT_genes %>% 
  filter(ensembl_gene_id %in% rownames(TCGA)) %>% 
  filter(regulated_by_methylation) %>% 
  pull(ensembl_gene_id)

not_met_genes <- CT_genes %>% 
  filter(ensembl_gene_id %in% rownames(TCGA)) %>% 
  filter(!regulated_by_methylation) %>% 
  pull(ensembl_gene_id)
TCGA <- TCGA[c(met_genes, not_met_genes), ]

binary_mat_methDep <- assay(TCGA[met_genes,]) >= 1
methDep_score <- colSums(binary_mat_methDep) / length(met_genes)
TCGA$methDep_score <- methDep_score

binary_mat_NonMethDep <- assay(TCGA[not_met_genes,]) >= 1
NonMethDep_score <- colSums(binary_mat_NonMethDep) / length(not_met_genes)
TCGA$NonMethDep_score <- NonMethDep_score

bilan <- rbind(as_tibble(colData(CCLE)) %>%
        mutate(source = "CCLE") %>%
        dplyr::select(source, methDep_score, NonMethDep_score),
      as_tibble(colData(TCGA)) %>%
        mutate(source = "TCGA") %>%
        dplyr::select(source, methDep_score, NonMethDep_score)) %>% 
  pivot_longer(names_to = "methDependency", values_to = "score", -source) %>% 
  mutate("methDependency" = ifelse(methDependency == "methDep_score", "MethDep", "Non-MethDep"))

bilan <- bilan %>% 
  mutate(group = case_when(score == 0 ~ "0",
                           score >= 0 & score < 0.05 ~ "<5",
                           score >= 0.05 & score < 0.10 ~ "5-10",
                           score >= 0.10 & score < 0.15 ~ "10-15",
                           score >= 0.15 & score < 0.20 ~ "15-20",
                           score >= 0.20 ~ ">20")) %>% 
  mutate(group = factor(group, levels = c("0", "<5", "5-10", "10-15","15-20", ">20")))
strip <- strip_themed(background_x = elem_list_rect(fill = c("gray95","gray80")))
fig <- bilan %>% 
  group_by(source, methDependency, group) %>% 
  summarize(n = n())  %>% #filter(methDependency == "MethDep") %>% 
  ggplot(aes(x = group, y = n, group = methDependency, fill = methDependency)) +
  geom_col(stat = "identity", width = 0.6, position = position_dodge(0.6))+
  scale_fill_manual(values = MethDep_colors)+
  xlab("Activation scores") +
  ylab("Number of tumors") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        #aspect.ratio = 1, 
        legend.position = c(0.65, 0.95),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 0),
        legend.key.size = unit(0.2, 'cm'),
        legend.background = element_blank()) +
  facet_wrap2(~source, ncol = 1, scales = "free_y", strip = strip)
fig
pdf(file = "../figs/Fig5D_ActivationScores.pdf", width = 3, height = 4)
fig
dev.off()
```